### 2022-12-02 writing/main.tex

\documentclass{article}
\usepackage{graphicx, setspace, booktabs, float, xcolor, rotating, lscape, lscape, sectsty, rotating, pdflscape}
\usepackage[utf8]{inputenc}
\usepackage[letterpaper, portrait, margin=0.8in]{geometry}
\usepackage[shortlabels]{enumitem}
\usepackage[round, sort]{natbib}

\doublespacing

\title{Thesis Updates}
\author{Arjun}
\date{December 2, 2022}

\begin{document}
\input{output/01_exploratory/tables/summary_statistics.tex}

\begin{landscape}
\input{output/01_exploratory/tables/balance_table.tex}
\end{landscape}

\end{document}