### 2022-12-07 output/01_exploratory/tables/bldg_val_results_table.tex
Created estimation tables

{
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\begin{tabular}{l*{4}{c}}
\toprule
                    &\multicolumn{1}{c}{\hspace{0.25cm}Building value}&\multicolumn{1}{c}{\hspace{0.25cm}Building value}&\multicolumn{1}{c}{\hspace{0.25cm}Building value}&\multicolumn{1}{c}{\hspace{0.25cm}Building value}\\
                    &        b/se&        b/se&        b/se&        b/se\\
\midrule
judgment\_for\_plaintiff& 2434984.290& 1161293.951& 1161293.951& 1055231.099\\
                    & (588147.83)& (453836.79)& (453836.79)&(2319051.94)\\
\midrule
Observations        &        5567&        5567&        5567&        5567\\
\bottomrule
\end{tabular}
}