### 2022-12-07 output/01_exploratory/tables/total_val_results_table.tex

{
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\begin{tabular}{l*{4}{c}}
\toprule
                    &\multicolumn{1}{c}{\hspace{0.25cm}Total property value}&\multicolumn{1}{c}{\hspace{0.25cm}Total property value}&\multicolumn{1}{c}{\hspace{0.25cm}Total property value}&\multicolumn{1}{c}{\hspace{0.25cm}Total property value}\\
                    &        b/se&        b/se&        b/se&        b/se\\
\midrule
judgment\_for\_plaintiff& 3259177.986& 1507062.672& 1507062.672&  751962.414\\
                    & (762388.53)& (577088.19)& (577088.19)&(2866558.93)\\
\midrule
Observations        &        5567&        5567&        5567&        5567\\
\bottomrule
\end{tabular}
}