### 2022-12-07 output/01_exploratory/tables/bldg_val_results_table.tex
Created estimation tables

{
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\begin{tabular}{l*{4}{c}}
\toprule
                    &\multicolumn{1}{c}{\hspace{0.25cm}Building value}&\multicolumn{1}{c}{\hspace{0.25cm}Building value}&\multicolumn{1}{c}{\hspace{0.25cm}Building value}&\multicolumn{1}{c}{\hspace{0.25cm}Building value}\\
                    &        b/se         &        b/se         &        b/se         &        b/se         \\
\midrule
judgment\_for\_plaintiff& 2434984.290\sym{***}& 1161293.951\sym{*}  & 1161293.951\sym{*}  & 1055231.099         \\
                    & (588147.83)         & (453836.79)         & (453836.79)         &(2319051.94)         \\
\midrule
Observations        &        5567         &        5567         &        5567         &        5567         \\
$\text{R}^2$        &     0.00307         &       0.491         &       0.491         &       0.527         \\
City F.E.           &          No         &         Yes         &         Yes         &          No         \\
Year F.E.           &          No         &         Yes         &         Yes         &          No         \\
Case Controls       &          No         &          No         &         Yes         &          No         \\
I.V. Estimate       &          No         &          No         &          No         &         Yes         \\
\bottomrule
\end{tabular}
}