### 2022-12-07 output/01_exploratory/tables/total_val_results_table.tex

{
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\begin{tabular}{l*{4}{c}}
\toprule
                    &\multicolumn{1}{c}{\hspace{0.25cm}Total property value}&\multicolumn{1}{c}{\hspace{0.25cm}Total property value}&\multicolumn{1}{c}{\hspace{0.25cm}Total property value}&\multicolumn{1}{c}{\hspace{0.25cm}Total property value}\\
                    &        b/se         &        b/se         &        b/se         &        b/se         \\
\midrule
judgment\_for\_plaintiff& 4783117.575\sym{***}& 2254653.696\sym{**} & 2254653.696\sym{**} & 3932513.092         \\
                    &(1111255.42)         & (832324.50)         & (832324.50)         &(4379643.76)         \\
\midrule
Observations        &        3331         &        3331         &        3331         &        3331         \\
$\text{R}^2$        &     0.00553         &       0.509         &       0.509         &       0.550         \\
City F.E.           &          No         &         Yes         &         Yes         &          No         \\
Year F.E.           &          No         &         Yes         &         Yes         &          No         \\
Case Controls       &          No         &          No         &         Yes         &          No         \\
I.V. Estimate       &          No         &          No         &          No         &         Yes         \\
\bottomrule
\end{tabular}
}