### 2022-12-07 output/01_exploratory/tables/ln_total_val_results_table.tex
Added placebo first stage regs and edited main results

\begin{table}[htbp]\centering
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\caption{Estimates of the Impact of Eviction}
\begin{tabular}{l*{3}{c}}
\toprule
                    &\multicolumn{1}{c}{ln\_total\_val}&\multicolumn{1}{c}{ln\_total\_val}&\multicolumn{1}{c}{ln\_total\_val}\\
\midrule
Plaintiff Victory   &       0.326\sym{***}&       0.171\sym{*}  &      -3.209         \\
                    &      (0.08)         &      (0.07)         &      (8.08)         \\
\midrule
Observations        &        3331         &        3331         &        2251         \\
$\text{R}^2$        &     0.00510         &       0.321         &       0.529         \\
City F.E.           &          No         &         Yes         &          No         \\
Year F.E.           &          No         &         Yes         &          No         \\
Controls for Potential I.V. Confounders&          No         &          No         &         Yes         \\
I.V. Estimate       &          No         &          No         &         Yes         \\
\bottomrule
\end{tabular}
\end{table}