### 2022-12-07 output/01_exploratory/tables/ln_total_val_results_table.tex
Removed iv confounders from regression

\begin{table}[htbp]\centering
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\caption{Estimates of the Impact of Eviction}
\begin{tabular}{l*{3}{c}}
\toprule
                    &\multicolumn{1}{c}{Log of total value}&\multicolumn{1}{c}{Log of total value}&\multicolumn{1}{c}{Log of total value}\\
\midrule
Plaintiff Victory   &       0.299\sym{***}&       0.142\sym{*}  &       0.189         \\
                    &      (0.08)         &      (0.07)         &      (0.33)         \\
\midrule
Observations        &        3699         &        3699         &        3699         \\
$\text{R}^2$        &     0.00420         &       0.349         &       0.358         \\
City F.E.           &          No         &         Yes         &          No         \\
Year F.E.           &          No         &         Yes         &          No         \\
Controls for Potential I.V. Confounders&          No         &          No         &         Yes         \\
I.V. Estimate       &          No         &          No         &         Yes         \\
\bottomrule
\end{tabular}
\end{table}