### 2022-12-07 writing/presentations/december_7.tex

\documentclass [xcolor=svgnames, t] {beamer} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs, comment} 
\usepackage[absolute, overlay]{textpos} 
\usepackage{pgfpages}
\usepackage[font=footnotesize]{caption}
\useoutertheme{infolines}

\definecolor{brownbrown}{RGB}{56, 28, 0}
\definecolor{brownred}{RGB}{228, 0, 43}

\setbeamercolor{title in head/foot}{bg=brownred, fg=brownbrown}
\setbeamercolor{author in head/foot}{bg=myuniversity}
\setbeamertemplate{page number in head/foot}{}
\usepackage{csquotes}

\usepackage{amsmath}
\usepackage[makeroom]{cancel}

\usepackage{textpos}

\usepackage{tikz}

\usetheme{Madrid}
\definecolor{myuniversity}{RGB}{56, 28, 0}
\usecolortheme[named=myuniversity]{structure}
\usepackage{tikz}

\title[Eviction and Property Values]{Estimating the Effect of Eviction on Assessment Values}

\institute[]{Department of Earth, Environmental, and Planetary Sciences  \\Brown University}

\author[Arjun Shanmugam]{}

\date{\today}

\addtobeamertemplate{navigation symbols}{}{%
    \usebeamerfont{footline}%
    \usebeamercolor[fg]{footline}%
    \hspace{1em}%
    \insertframenumber/\inserttotalframenumber
}

\begin{document}
\begin{frame}
\maketitle
\end{frame}

\begin{frame}
\frametitle{Table of Contents}
\tableofcontents
\end{frame}

\section{Introduction}
\begin{frame}{Introduction and Literature}
   \begin{enumerate}
       \item Eviction is a bigger problem in America than in other rich countries
       \begin{enumerate}
           \item Most poor renting families in America
spend over half of their incomes on rent (Desmond, 2017)
       \end{enumerate}
       \item Sociologists have found large effects of evictions using non-experimental methods
       \begin{enumerate}
           \item Increased probability of job loss (Desmond, Gershenson, 2016)
           \item Increased probability of depression (Desmond, Kimbro, 2015)
       \end{enumerate} 
       \item However, these estimates are likely confounded
       \begin{enumerate}
           \item ``Ashenfelter dip''
           \item Observable and unobservable confounders
       \end{enumerate}
       \item Only recently have economists began to apply quasi-experimental methods to study eviction
        \begin{enumerate}
            \item Van Dijk et al. find effects on durable consumption, access to credit, and more
        \end{enumerate}
   \end{enumerate}
\end{frame}

\section{Motivation}

\begin{frame}{Motivation}
\begin{enumerate}
    \item Strong negative correlation between the presence of eviction and property values
    \item Several potential channels through which eviction might affect property values
    \begin{enumerate}
        \item Evidence suggests that longer tenancy can increase property values
        \item Or, perhaps landlords are evicting problematic tenants
    \end{enumerate}
    \item Either way—property values determine school district 
    funding, so kids' outcomes could be at stake
    \item Ideally, I would have \emph{market} values instead of assessment values
\end{enumerate}
\end{frame}

\section{Data}
\begin{frame}{Data}
    \begin{enumerate}
        \item Eviction data from MassLandlords, a trade association
        \begin{enumerate}
            \item Scraped around 50,000 eviction case records
            \item Collects case outcome data, including method of resolution, judge name, and more
        \end{enumerate}
        \item Assessment values data from the Massachusetts Office of Geographic Information Services
        \begin{enumerate}
            \item Unbalanced panel data on assessor values for nearly every property in Massachussetts
        \end{enumerate}
        \item Link evictions to assessment records on fiscal year and location
    \end{enumerate}
\end{frame}

\begin{frame}{Data}
\begin{table}[H]
    \centering
    \tiny
    \input{output/01_exploratory/tables/summary_statistics.tex}
    \caption{Caption}
    \label{tab:my_label}
\end{table}
\end{frame}

\input{output/01_exploratory/tables/balance_table.tex}

\section{Empirical Strategy}

\begin{frame} [allowframebreaks]\frametitle{References}

        \bibliographystyle{apalike}
        \bibliography{bibfile}
\end{frame}

\end{document}
Close