### 2022-12-20 output/01_exploratory/tables/ln_total_val_results_table.tex
Ran DiD

\begin{table}[htbp]\centering
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\caption{Estimates of the Impact of Eviction}
\begin{tabular}{l*{3}{c}}
\toprule
                    &\multicolumn{1}{c}{Log(Total Value)}&\multicolumn{1}{c}{Log(Total Value)}&\multicolumn{1}{c}{Log(Total Value)}\\
\midrule
Plaintiff Victory   &       0.283\sym{***}&       0.142\sym{*}  &       0.267         \\
                    &      (0.07)         &      (0.06)         &      (0.32)         \\
\midrule
Observations        &        4115         &        4115         &        4115         \\
$\text{R}^2$        &     0.00378         &       0.376         &       0.394         \\
City F.E.           &          No         &         Yes         &          No         \\
Year F.E.           &          No         &         Yes         &          No         \\
I.V. Estimate       &          No         &          No         &         Yes         \\
\bottomrule
\end{tabular}
\end{table}