### 2022-12-25 output/summary_statistics/tables/balance_table.tex
Added GETting of all zpids when several are returned for one address

\begin{tabular}{llcccc}
\toprule
 &  & \multicolumn{2}{c}{\textit{Cases Won by Defendant}} & \multicolumn{2}{c}{\textit{Cases Won by Plaintiff}} \\
 &  & Mean & N & Mean & N \\
Panel & Variable &  &  &  &  \\
\midrule
\multirow[c]{8}{4cm}{\textit{Panel A: Case Initiation}} & For cause & 0.12 & 4090 & 0.12 & 11921 \\
 & For cause (transfer) & 0.01 & 4090 & 0.01 & 11921 \\
 & Foreclosure & 0.01 & 4090 & 0.03 & 11921 \\
 & Foreclosure (transfer) & 0.00 & 4090 & 0.00 & 11921 \\
 & No cause & 0.12 & 4090 & 0.10 & 11921 \\
 & No cause (transfer) & 0.01 & 4090 & 0.00 & 11921 \\
 & Non-payment of rent & 0.69 & 4090 & 0.72 & 11921 \\
 & Non-payment of rent (transfer) & 0.03 & 4090 & 0.01 & 11921 \\
\cline{1-6}
\multirow[c]{4}{4cm}{\textit{Panel B: Case Resolution}} & Case defaulted & 0.00 & 4090 & 0.67 & 11921 \\
 & Case involuntarily dismissed & 0.99 & 4090 & 0.00 & 11921 \\
 & Case heard & 0.01 & 4090 & 0.19 & 11921 \\
 & Case mediated & 0.00 & 4090 & 0.00 & 11921 \\
\cline{1-6}
\multirow[c]{5}{4cm}{\textit{Panel C: Defendant and Plaintiff Characteristics}} & Defendant has an attorney & 0.20 & 4090 & 0.07 & 11921 \\
 & Plaintiff has an attorney & 0.79 & 4090 & 0.81 & 11921 \\
 & Defendant is an entity & 0.01 & 4090 & 0.01 & 11921 \\
 & Plaintiff is an entity & 0.63 & 4090 & 0.67 & 11921 \\
 & Money judgment & 57.56 & 4090 & 3444.31 & 11921 \\
\cline{1-6}
\multirow[c]{5}{4cm}{\textit{Panel D: Assessor Records From Post-Filing F.Y.}} & Total property value & 6478531.48 & 1256 & 8995455.63 & 3992 \\
 & Building value & 1605427.57 & 1256 & 2352344.15 & 3992 \\
 & Land value & 61247.49 & 1256 & 100993.11 & 3992 \\
 & Other value & 8138231.73 & 1256 & 11437647.57 & 3992 \\
 & Units & 41.23 & 1256 & 42.09 & 3992 \\
\cline{1-6}
\multirow[c]{9}{4cm}{\textit{Panel E: Zestimates Around Last Docket Date}} & Five years before latest docket date & 245078.62 & 957 & 275426.28 & 2603 \\
 & Four years before latest docket date & 265248.64 & 977 & 294572.61 & 2635 \\
 & Three years before latest docket date & 308529.26 & 992 & 317382.82 & 2706 \\
 & Two years before latest docket date & 330329.21 & 998 & 346565.14 & 2709 \\
 & One year before latest docket date & 336544.68 & 1002 & 364931.93 & 2715 \\
 & Latest docket date & 364629.48 & 991 & 391919.00 & 2701 \\
 & One year after latest docket date & 412421.05 & 1045 & 440748.11 & 2796 \\
 & Two years after latest docket date & 450898.44 & 801 & 509321.48 & 2493 \\
 & Three years after latest docket date & 485688.38 & 441 & 512215.31 & 1613 \\
\cline{1-6}
\bottomrule
\end{tabular}