### 2022-12-28 output/summary_statistics/tables/balance_table.tex
About to revamp DiD observed periodS

\begin{tabular}{llcccc}
\toprule
 &  & \multicolumn{2}{c}{\textit{Cases Won by Defendant}} & \multicolumn{2}{c}{\textit{Cases Won by Plaintiff}} \\
 &  & Mean & N & Mean & N \\
Panel & Variable &  &  &  &  \\
\midrule
\multirow[c]{8}{4cm}{\textit{Panel A: Case Initiation}} & For cause & 0.12 & 4090 & 0.12 & 11921 \\
 & For cause (transfer) & 0.01 & 4090 & 0.01 & 11921 \\
 & Foreclosure & 0.01 & 4090 & 0.03 & 11921 \\
 & Foreclosure (transfer) & 0.00 & 4090 & 0.00 & 11921 \\
 & No cause & 0.12 & 4090 & 0.10 & 11921 \\
 & No cause (transfer) & 0.01 & 4090 & 0.00 & 11921 \\
 & Non-payment of rent & 0.69 & 4090 & 0.72 & 11921 \\
 & Non-payment of rent (transfer) & 0.03 & 4090 & 0.01 & 11921 \\
\cline{1-6}
\multirow[c]{4}{4cm}{\textit{Panel B: Case Resolution}} & Case defaulted & 0.00 & 4090 & 0.67 & 11921 \\
 & Case involuntarily dismissed & 0.99 & 4090 & 0.00 & 11921 \\
 & Case heard & 0.01 & 4090 & 0.19 & 11921 \\
 & Case mediated & 0.00 & 4090 & 0.00 & 11921 \\
\cline{1-6}
\multirow[c]{5}{4cm}{\textit{Panel C: Defendant and Plaintiff Characteristics}} & Defendant has an attorney & 0.20 & 4090 & 0.07 & 11921 \\
 & Plaintiff has an attorney & 0.79 & 4090 & 0.81 & 11921 \\
 & Defendant is an entity & 0.01 & 4090 & 0.01 & 11921 \\
 & Plaintiff is an entity & 0.63 & 4090 & 0.67 & 11921 \\
 & Money judgment & 57.56 & 4090 & 3444.31 & 11921 \\
\cline{1-6}
\multirow[c]{5}{4cm}{\textit{Panel D: Assessor Records From Post-Filing F.Y.}} & Total property value & 6478531.48 & 1256 & 8995455.63 & 3992 \\
 & Building value & 1605427.57 & 1256 & 2352344.15 & 3992 \\
 & Land value & 61247.49 & 1256 & 100993.11 & 3992 \\
 & Other value & 8138231.73 & 1256 & 11437647.57 & 3992 \\
 & Units & 41.23 & 1256 & 42.09 & 3992 \\
\cline{1-6}
\multirow[c]{9}{4cm}{\textit{Panel E: Zestimates Around Last Docket Date}} & Five years before latest docket date & 255601.88 & 1019 & 267829.81 & 2904 \\
 & Four years before latest docket date & 278440.18 & 1035 & 286996.41 & 2939 \\
 & Three years before latest docket date & 297578.04 & 1052 & 313182.84 & 3015 \\
 & Two years before latest docket date & 332743.89 & 1055 & 341145.74 & 3016 \\
 & One year before latest docket date & 338598.45 & 1058 & 360142.85 & 3018 \\
 & Latest docket date & 373816.17 & 1044 & 387035.91 & 2993 \\
 & One year after latest docket date & 419391.34 & 1086 & 431642.67 & 3087 \\
 & Two years after latest docket date & 446124.50 & 824 & 491836.71 & 2726 \\
 & Three years after latest docket date & 482504.08 & 455 & 508987.34 & 1793 \\
\cline{1-6}
\bottomrule
\end{tabular}