### 2022-12-31 output/summary_statistics/tables/balance_table.tex
Added median to summary statistics

\begin{tabular}{llcccc}
\toprule
 &  & \multicolumn{2}{c}{\textit{Cases Won by Defendant}} & \multicolumn{2}{c}{\textit{Cases Won by Plaintiff}} \\
 &  & Mean & N & Mean & N \\
Panel & Variable &  &  &  &  \\
\midrule
\multirow[c]{8}{4cm}{\textit{Panel A: Case Initiation}} & For cause & 0.12 & 4082 & 0.12 & 11912 \\
 & For cause (transfer) & 0.01 & 4082 & 0.01 & 11912 \\
 & Foreclosure & 0.01 & 4082 & 0.03 & 11912 \\
 & Foreclosure (transfer) & 0.00 & 4082 & 0.00 & 11912 \\
 & No cause & 0.12 & 4082 & 0.10 & 11912 \\
 & No cause (transfer) & 0.01 & 4082 & 0.00 & 11912 \\
 & Non-payment of rent & 0.69 & 4082 & 0.72 & 11912 \\
 & Non-payment of rent (transfer) & 0.03 & 4082 & 0.01 & 11912 \\
\cline{1-6}
\multirow[c]{5}{4cm}{\textit{Panel B: Case Resolution}} & Case duration & 97.63 & 4082 & 58.85 & 10273 \\
 & Case defaulted & 0.00 & 4082 & 0.67 & 11912 \\
 & Case involuntarily dismissed & 0.99 & 4082 & 0.00 & 11912 \\
 & Case heard & 0.01 & 4082 & 0.19 & 11912 \\
 & Case mediated & 0.00 & 4082 & 0.00 & 11912 \\
\cline{1-6}
\multirow[c]{5}{4cm}{\textit{Panel C: Defendant and Plaintiff Characteristics}} & Defendant has an attorney & 0.20 & 4082 & 0.07 & 11912 \\
 & Plaintiff has an attorney & 0.79 & 4082 & 0.81 & 11912 \\
 & Defendant is an entity & 0.01 & 4082 & 0.01 & 11912 \\
 & Plaintiff is an entity & 0.63 & 4082 & 0.67 & 11912 \\
 & Money judgment & 57.67 & 4082 & 3440.73 & 11912 \\
\cline{1-6}
\multirow[c]{5}{4cm}{\textit{Panel D: Assessor Records From Post-Filing F.Y.}} & Total property value & 6478531.48 & 1256 & 8997593.33 & 3991 \\
 & Building value & 1605427.57 & 1256 & 2352903.60 & 3991 \\
 & Land value & 61247.49 & 1256 & 101018.42 & 3991 \\
 & Other value & 8138231.73 & 1256 & 11440367.23 & 3991 \\
 & Units & 41.23 & 1256 & 42.10 & 3991 \\
\cline{1-6}
\multirow[c]{9}{4cm}{\textit{Panel E: Zestimates Around Last Docket Date}} & Five years before latest docket date & 255719.80 & 1018 & 267442.86 & 2901 \\
 & Four years before latest docket date & 278556.70 & 1034 & 285983.57 & 2936 \\
 & Three years before latest docket date & 297696.16 & 1051 & 312711.10 & 3012 \\
 & Two years before latest docket date & 332887.98 & 1054 & 339939.67 & 3013 \\
 & One year before latest docket date & 338734.41 & 1057 & 359532.30 & 3015 \\
 & Latest docket date & 373978.71 & 1043 & 385699.20 & 2990 \\
 & One year after latest docket date & 419565.10 & 1085 & 431037.20 & 3084 \\
 & Two years after latest docket date & 446321.14 & 823 & 490315.74 & 2723 \\
 & Three years after latest docket date & 482787.35 & 454 & 507050.53 & 1791 \\
\cline{1-6}
\bottomrule
\end{tabular}