### 2023-01-17 output/DiD/tables/balance_table.tex
Removed vertical lines around outer level of columns in balance table

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & Zestimate, Jan. 2017 & 291,942.14 & 26,570.00 & 0.00 & 10,030.65 & 0.18 \\
 & Zestimate, Jan. 2019 $-$ Zestimate, Jan. 2017 & 50,491.60 & 2,664.69 & 0.27 & -23.86 & 0.99 \\
\cline{1-7}
\multirow[c]{6}{3cm}{\textit{Panel B}} & Share with bachelor's degree & 0.25 & 0.01 & 0.07 & 0.01 & 0.40 \\
 & Jobs per sq. mile (2010) & 4,158.02 & 263.90 & 0.63 & 122.36 & 0.83 \\
 & Median household income (2016) & 52,087.32 & 2,566.82 & 0.01 & 1,159.98 & 0.25 \\
 & Population density (2010) & 9,031.64 & 944.99 & 0.01 & 488.36 & 0.17 \\
 & Share white (2010) & 0.60 & 0.04 & 0.00 & 0.02 & 0.06 \\
 & Share with commute <15 min. (2010) & 0.29 & -0.01 & 0.02 & -0.00 & 0.27 \\
\cline{1-7}
\textit{Panel C} & For cause & 0.12 & 0.01 & 0.68 & 0.00 & 0.83 \\
\cline{1-7}
\textit{Panel D} & Defendant is an entity & 0.01 & 0.00 & 0.37 & 0.00 & 0.66 \\
\cline{1-7}
\bottomrule
\end{tabular}