### 2023-01-19 output/DiD/tables/balance_table.tex
Dropped months w low sample size

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & Zestimate, Jan. 2017 & 291,942.14 & 26,566.34 & 0.00 & 14,752.11 & 0.14 \\
 & Change, Jan. 2017 to Jan. 2019 & 50,491.60 & 2,654.65 & 0.27 & -2,202.60 & 0.49 \\
\cline{1-7}
\multirow[c]{7}{3cm}{\textit{Panel B}} & Share with bachelor's degree & 0.25 & 0.01 & 0.07 & 0.00 & 0.96 \\
 & Jobs per square mile (2010) & 4,158.02 & 263.90 & 0.63 & 322.24 & 0.64 \\
 & Median household income (2016) & 52,087.32 & 2,566.82 & 0.01 & 1,497.45 & 0.24 \\
 & Population density (2010) & 9,031.64 & 944.99 & 0.01 & 787.73 & 0.09 \\
 & Median two bedroom rent (2015) & 1,093.36 & 34.26 & 0.05 & 6.43 & 0.77 \\
 & Share white (2010) & 0.60 & 0.04 & 0.00 & 0.03 & 0.02 \\
 & Share with commute $<$15 minutes (2010) & 0.29 & -0.01 & 0.02 & -0.00 & 0.60 \\
\cline{1-7}
\textit{Panel C} & For cause & 0.12 & 0.01 & 0.68 & -0.00 & 0.86 \\
\cline{1-7}
\textit{Panel D} & Defendant is an entity & 0.01 & 0.00 & 0.37 & 0.00 & 0.89 \\
\cline{1-7}
\bottomrule
\end{tabular}