### 2023-01-19 output/DiD/tables/pre_treatment_covariate_tests.tex

\begin{tabular}{llcc}
\toprule
 &  & \multicolumn{2}{c}{\textit{Dependent Variable}} \\
\cline{3-4}
\\
 &  & Zestimate, Dec. 2022 & Plaintiff victory \\
 & \emph{Independent Variable} &  &  \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A: Pre-treatment Zestimates}} & Zestimate, Jan. 2017 & 0.00 & 0.00 \\
 & Change, Jan. 2017 to Jan. 2019 & 0.00 & 0.27 \\
\cline{1-4}
\multirow[c]{8}{3cm}{\textit{Panel B: Census Tract Characteristics}} & Share with bachelor's degree & 0.00 & 0.07 \\
 & Jobs per square mile (2010) & 0.00 & 0.63 \\
 & Median household income (2016) & 0.00 & 0.01 \\
 & Share below poverty line & 0.70 & 0.01 \\
 & Population density (2010) & 0.00 & 0.01 \\
 & Median two bedroom rent (2015) & 0.00 & 0.05 \\
 & Share white (2010) & 0.00 & 0.00 \\
 & Share with commute $<$15 minutes (2010) & 0.00 & 0.02 \\
\cline{1-4}
\multirow[c]{3}{3cm}{\textit{Panel C: Case Initiation}} & For cause & 0.01 & 0.68 \\
 & No cause & 0.98 & 0.03 \\
 & Non-payment of rent & 0.31 & 0.77 \\
\cline{1-4}
\multirow[c]{4}{3cm}{\textit{Panel D: Defendant and Plaintiff Characteristics}} & Defendant has an attorney & 0.37 & 0.00 \\
 & Plaintiff has an attorney & 0.61 & 0.47 \\
 & Defendant is an entity & 0.02 & 0.37 \\
 & Plaintiff is an entity & 0.62 & 0.13 \\
\cline{1-4}
\bottomrule
\end{tabular}