### 2023-02-25 output/DiD/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & Zestimate, Jan. 2017 & 291,945.81 & 26,673.96 & 0.00 & 3,189.17 & 0.75 \\
 & Change, Jan. 2017 to Jan. 2019 & 50,501.64 & 2,697.85 & 0.27 & -3,196.88 & 0.32 \\
\cline{1-7}
\multirow[c]{7}{3cm}{\textit{Panel B}} & Share with bachelor's degree & 0.25 & 0.01 & 0.09 & -0.00 & 0.96 \\
 & Jobs per square mile (2010) & 4,161.51 & 259.78 & 0.64 & 668.27 & 0.37 \\
 & Median household income (2016) & 52,084.78 & 2,271.04 & 0.03 & -211.14 & 0.87 \\
 & Population density (2010) & 9,018.27 & 962.56 & 0.01 & 644.23 & 0.17 \\
 & Median two bedroom rent (2015) & 1,093.52 & 33.19 & 0.05 & 8.19 & 0.69 \\
 & Share white (2010) & 0.60 & 0.03 & 0.00 & -0.00 & 0.94 \\
 & Share with commute $<$15 minutes (2010) & 0.29 & -0.01 & 0.02 & -0.01 & 0.29 \\
\cline{1-7}
\textit{Panel C} & For cause & 0.12 & 0.01 & 0.67 & -0.01 & 0.56 \\
\cline{1-7}
\textit{Panel D} & Defendant is an entity & 0.01 & 0.00 & 0.36 & 0.00 & 0.64 \\
\cline{1-7}
\bottomrule
\end{tabular}