### 2023-02-27 output/DiD_crime/tables/balance_table.tex
Abstracted DiD code

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & Crimes reported, Jan. 2017 & 1.51 & 0.35 & 0.57 & 0.14 & 0.82 \\
 & Change in Crimes Reported, Jan. 2017 to Jan. 2019 & -0.02 & 0.07 & 0.50 & 0.05 & 0.65 \\
\cline{1-7}
\textit{Panel B} & Share with commute $<$15 minutes (2010) & 0.16 & 0.01 & 0.23 & -0.01 & 0.11 \\
\cline{1-7}
\textit{Panel D} & Plaintiff is an entity & 0.81 & 0.10 & 0.00 & 0.01 & 0.16 \\
\cline{1-7}
\bottomrule
\end{tabular}