### 2023-03-08 output/crime_140m/tables/balance_table.tex
ra models

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & twenty_seventeen_any_crime_140m & 9.70 & 0.56 & 0.31 & 1.64 & 0.13 \\
 & pre_treatment_change_in_any_crime_140m & -1.03 & 0.02 & 0.95 & -0.12 & 0.86 \\
\cline{1-7}
\multirow[c]{7}{3cm}{\textit{Panel B}} & Share with bachelor's degree & 0.32 & 0.01 & 0.24 & 0.07 & 0.00 \\
 & Jobs per square mile (2010) & 16,069.14 & 3,197.93 & 0.16 & 12,479.47 & 0.00 \\
 & Median household income (2016) & 48,035.65 & 1,777.65 & 0.19 & 3,359.04 & 0.19 \\
 & Population density (2010) & 23,086.04 & 1,373.15 & 0.06 & 511.85 & 0.72 \\
 & Median two bedroom rent (2015) & 1,339.99 & -29.55 & 0.50 & 177.05 & 0.03 \\
 & Share white (2010) & 0.31 & 0.02 & 0.10 & 0.09 & 0.00 \\
 & Share with commute $<$15 minutes (2010) & 0.16 & 0.01 & 0.23 & 0.02 & 0.02 \\
\cline{1-7}
\multirow[c]{2}{3cm}{\textit{Panel C}} & No cause & 0.05 & -0.05 & 0.00 & -0.11 & 0.00 \\
 & Non-payment of rent & 0.78 & 0.07 & 0.00 & 0.16 & 0.00 \\
\cline{1-7}
\multirow[c]{2}{3cm}{\textit{Panel D}} & Plaintiff has an attorney & 0.88 & 0.07 & 0.00 & 0.21 & 0.00 \\
 & Plaintiff is an entity & 0.81 & 0.10 & 0.00 & 0.23 & 0.00 \\
\cline{1-7}
\bottomrule
\end{tabular}