### 2023-03-08 output/crime_140m/tables/pre_treatment_covariate_tests.tex

\begin{tabular}{llcc}
\toprule
 &  & \multicolumn{2}{c}{\textit{Dependent Variable}} \\
\cline{3-4}
\\
 &  & Any Crime Incidents Within 140m, Dec. 2022 & Plaintiff Victory \\
 & \emph{Independent Variable} &  &  \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A: Pre-treatment Outcomes}} & twenty_seventeen_any_crime_140m & 0.00 & 0.31 \\
 & pre_treatment_change_in_any_crime_140m & 0.00 & 0.95 \\
\cline{1-4}
\multirow[c]{8}{3cm}{\textit{Panel B: Census Tract Characteristics}} & Share with bachelor's degree & 0.00 & 0.24 \\
 & Jobs per square mile (2010) & 0.00 & 0.16 \\
 & Median household income (2016) & 0.00 & 0.19 \\
 & Share below poverty line & 0.09 & 0.14 \\
 & Population density (2010) & 0.00 & 0.06 \\
 & Median two bedroom rent (2015) & 0.00 & 0.50 \\
 & Share white (2010) & 0.00 & 0.10 \\
 & Share with commute $<$15 minutes (2010) & 0.00 & 0.23 \\
\cline{1-4}
\multirow[c]{3}{3cm}{\textit{Panel C: Case Initiation}} & For cause & 0.18 & 0.48 \\
 & No cause & 0.03 & 0.00 \\
 & Non-payment of rent & 0.01 & 0.00 \\
\cline{1-4}
\multirow[c]{4}{3cm}{\textit{Panel D: Defendant and Plaintiff Characteristics}} & Defendant has an attorney & 0.09 & 0.00 \\
 & Plaintiff has an attorney & 0.00 & 0.00 \\
 & Defendant is an entity & 0.86 & 0.00 \\
 & Plaintiff is an entity & 0.00 & 0.00 \\
\cline{1-4}
\bottomrule
\end{tabular}