### 2023-03-08 output/crime_60m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\textit{Panel A} & twenty_seventeen_any_crime_60m & 2.03 & -0.04 & 0.83 & 0.12 & 0.66 \\
\cline{1-7}
\multirow[c]{5}{3cm}{\textit{Panel B}} & Share with bachelor's degree & 0.32 & 0.01 & 0.24 & 0.07 & 0.00 \\
 & Jobs per square mile (2010) & 16,069.14 & 3,197.93 & 0.16 & 12,355.85 & 0.00 \\
 & Population density (2010) & 23,086.04 & 1,373.15 & 0.06 & 533.46 & 0.71 \\
 & Median two bedroom rent (2015) & 1,339.99 & -29.55 & 0.50 & 174.33 & 0.03 \\
 & Share with commute $<$15 minutes (2010) & 0.16 & 0.01 & 0.23 & 0.02 & 0.02 \\
\cline{1-7}
\textit{Panel C} & Non-payment of rent & 0.78 & 0.07 & 0.00 & 0.16 & 0.00 \\
\cline{1-7}
\multirow[c]{3}{3cm}{\textit{Panel D}} & Defendant has an attorney & 0.06 & -0.18 & 0.00 & -0.23 & 0.00 \\
 & Plaintiff has an attorney & 0.88 & 0.07 & 0.00 & 0.21 & 0.00 \\
 & Plaintiff is an entity & 0.81 & 0.10 & 0.00 & 0.23 & 0.00 \\
\cline{1-7}
\bottomrule
\end{tabular}