### 2023-03-12 output/crime_280m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & twenty_seventeen_any_crime_280m & 33.56 & 0.76 & 0.57 & 0.72 & 0.77 \\
 & pre_treatment_change_in_any_crime_280m & 0.54 & 1.55 & 0.01 & 0.33 & 0.78 \\
\cline{1-7}
\multirow[c]{7}{3cm}{\textit{Panel B}} & Jobs per square mile (2010) & 16,069.14 & 3,197.93 & 0.16 & 12,430.78 & 0.00 \\
 & Median household income (2016) & 48,035.65 & 1,777.65 & 0.19 & 3,350.72 & 0.19 \\
 & Share below poverty line & 0.27 & 0.01 & 0.14 & 0.00 & 0.86 \\
 & Population density (2010) & 23,086.04 & 1,373.15 & 0.06 & 516.07 & 0.71 \\
 & Median two bedroom rent (2015) & 1,339.99 & -29.55 & 0.50 & 175.33 & 0.03 \\
 & Share white (2010) & 0.31 & 0.02 & 0.10 & 0.09 & 0.00 \\
 & Share with commute $<$15 minutes (2010) & 0.16 & 0.01 & 0.23 & 0.02 & 0.02 \\
\cline{1-7}
\textit{Panel C} & No cause & 0.05 & -0.05 & 0.00 & -0.11 & 0.00 \\
\cline{1-7}
\textit{Panel D} & Plaintiff is an entity & 0.81 & 0.10 & 0.00 & 0.23 & 0.00 \\
\cline{1-7}
\bottomrule
\end{tabular}