### 2023-03-12 output/crime_60m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & twenty_seventeen_any_crime_60m & 0.47 & 0.03 & 0.49 & -0.00 & 0.95 \\
 & pre_treatment_change_in_any_crime_60m & -0.03 & 0.01 & 0.79 & 0.01 & 0.87 \\
\cline{1-7}
\multirow[c]{3}{3cm}{\textit{Panel B}} & Jobs per square mile (2010) & 16,069.14 & 3,197.93 & 0.16 & 903.79 & 0.69 \\
 & Population density (2010) & 23,086.04 & 1,373.15 & 0.06 & -271.38 & 0.68 \\
 & Share with commute $<$15 minutes (2010) & 0.16 & 0.01 & 0.23 & -0.00 & 0.42 \\
\cline{1-7}
\multirow[c]{2}{3cm}{\textit{Panel D}} & Plaintiff has an attorney & 0.88 & 0.07 & 0.00 & -0.00 & 0.93 \\
 & Plaintiff is an entity & 0.81 & 0.10 & 0.00 & 0.02 & 0.27 \\
\cline{1-7}
\bottomrule
\end{tabular}