### 2023-03-12 output/crime_90m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & twenty_seventeen_any_crime_90m & 0.68 & 0.02 & 0.77 & -0.12 & 0.26 \\
 & pre_treatment_change_in_any_crime_90m & -0.02 & 0.01 & 0.91 & 0.03 & 0.76 \\
\cline{1-7}
\multirow[c]{8}{3cm}{\textit{Panel B}} & Share with bachelor's degree & 0.32 & 0.01 & 0.24 & 0.07 & 0.00 \\
 & Jobs per square mile (2010) & 16,069.14 & 3,197.93 & 0.16 & 12,393.10 & 0.00 \\
 & Median household income (2016) & 48,035.65 & 1,777.65 & 0.19 & 3,315.55 & 0.20 \\
 & Share below poverty line & 0.27 & 0.01 & 0.14 & 0.00 & 0.85 \\
 & Population density (2010) & 23,086.04 & 1,373.15 & 0.06 & 537.03 & 0.70 \\
 & Median two bedroom rent (2015) & 1,339.99 & -29.55 & 0.50 & 177.10 & 0.03 \\
 & Share white (2010) & 0.31 & 0.02 & 0.10 & 0.09 & 0.00 \\
 & Share with commute $<$15 minutes (2010) & 0.16 & 0.01 & 0.23 & 0.02 & 0.02 \\
\cline{1-7}
\multirow[c]{2}{3cm}{\textit{Panel D}} & Plaintiff has an attorney & 0.88 & 0.07 & 0.00 & 0.21 & 0.00 \\
 & Plaintiff is an entity & 0.81 & 0.10 & 0.00 & 0.23 & 0.00 \\
\cline{1-7}
\bottomrule
\end{tabular}