### 2023-03-13 output/crime_140m/tables/balance_table.tex
updated analysis utilities to accomodate new outcomes

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\textit{Panel A} & twenty_seventeen_group_1_crimes_200m & 1.38 & 0.09 & 0.35 & 0.02 & 0.89 \\
\cline{1-7}
\multirow[c]{7}{3cm}{\textit{Panel B}} & frac_coll_plus2010 & 0.32 & 0.01 & 0.24 & 0.05 & 0.01 \\
 & job_density_2013 & 16,069.14 & 3,197.93 & 0.16 & 5,413.87 & 0.16 \\
 & poor_share2010 & 0.27 & 0.01 & 0.14 & -0.01 & 0.33 \\
 & popdensity2010 & 23,086.04 & 1,373.15 & 0.06 & 483.47 & 0.68 \\
 & rent_twobed2015 & 1,339.99 & -29.55 & 0.50 & 30.88 & 0.67 \\
 & share_white2010 & 0.31 & 0.02 & 0.10 & 0.05 & 0.04 \\
 & traveltime15_2010 & 0.16 & 0.01 & 0.23 & 0.01 & 0.20 \\
\cline{1-7}
\textit{Panel C} & no_cause & 0.05 & -0.05 & 0.00 & -0.08 & 0.00 \\
\cline{1-7}
\multirow[c]{3}{3cm}{\textit{Panel D}} & hasAttyP & 0.88 & 0.07 & 0.00 & 0.11 & 0.00 \\
 & isEntityD & 0.01 & -0.04 & 0.00 & 0.01 & 0.17 \\
 & isEntityP & 0.81 & 0.10 & 0.00 & 0.12 & 0.00 \\
\cline{1-7}
\bottomrule
\end{tabular}