### 2023-03-13 output/crime_200m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\textit{Panel A} & twenty_seventeen_group_0_crimes_50m & 1.43 & -0.20 & 0.16 & -0.16 & 0.26 \\
\cline{1-7}
\multirow[c]{3}{3cm}{\textit{Panel B}} & job_density_2013 & 16,069.14 & 3,197.93 & 0.16 & 843.09 & 0.71 \\
 & popdensity2010 & 23,086.04 & 1,373.15 & 0.06 & -365.58 & 0.58 \\
 & traveltime15_2010 & 0.16 & 0.01 & 0.23 & -0.00 & 0.34 \\
\cline{1-7}
\textit{Panel C} & non_payment & 0.78 & 0.07 & 0.00 & -0.00 & 0.92 \\
\cline{1-7}
\textit{Panel D} & isEntityP & 0.81 & 0.10 & 0.00 & 0.01 & 0.40 \\
\cline{1-7}
\bottomrule
\end{tabular}