### 2023-03-13 output/crime_90m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\textit{Panel A} & twenty_seventeen_group_1_crimes_150m & 0.86 & 0.14 & 0.05 & 0.03 & 0.68 \\
\cline{1-7}
\multirow[c]{4}{3cm}{\textit{Panel B}} & frac_coll_plus2010 & 0.32 & 0.01 & 0.24 & -0.01 & 0.60 \\
 & poor_share2010 & 0.27 & 0.01 & 0.14 & -0.01 & 0.43 \\
 & share_white2010 & 0.31 & 0.02 & 0.10 & -0.00 & 0.94 \\
 & traveltime15_2010 & 0.16 & 0.01 & 0.23 & -0.00 & 0.44 \\
\cline{1-7}
\textit{Panel C} & non_payment & 0.78 & 0.07 & 0.00 & 0.00 & 0.99 \\
\cline{1-7}
\textit{Panel D} & isEntityP & 0.81 & 0.10 & 0.00 & 0.01 & 0.38 \\
\cline{1-7}
\bottomrule
\end{tabular}