### 2023-03-15 output/summary_statistics/tables/summary_statistics.tex

\begin{tabular}{llcccc}
\toprule
 &  & Mean & Median & S.D. & N \\
Panel & Variable &  &  &  &  \\
\midrule
\multirow[c]{34}{4cm}{\textit{Panel A: Pre-treatment Outcomes}} & pre_treatment_change_in_group_0_crimes_100m & -0.08 & 0.00 & 1.65 & 40,734 \\
 & pre_treatment_change_in_group_0_crimes_150m & -0.19 & 0.00 & 2.74 & 40,734 \\
 & pre_treatment_change_in_group_0_crimes_200m & -0.36 & 0.00 & 3.68 & 40,734 \\
 & pre_treatment_change_in_group_0_crimes_50m & -0.01 & 0.00 & 0.82 & 40,734 \\
 & pre_treatment_change_in_group_1_crimes_100m & 0.00 & 0.00 & 0.39 & 40,734 \\
 & pre_treatment_change_in_group_1_crimes_150m & 0.01 & 0.00 & 0.53 & 40,734 \\
 & pre_treatment_change_in_group_1_crimes_200m & 0.02 & 0.00 & 0.69 & 40,734 \\
 & pre_treatment_change_in_group_1_crimes_50m & 0.00 & 0.00 & 0.21 & 40,734 \\
 & pre_treatment_change_in_group_2_crimes_100m & -0.00 & 0.00 & 0.37 & 40,734 \\
 & pre_treatment_change_in_group_2_crimes_150m & -0.01 & 0.00 & 0.59 & 40,734 \\
 & pre_treatment_change_in_group_2_crimes_200m & -0.03 & 0.00 & 0.84 & 40,734 \\
 & pre_treatment_change_in_group_2_crimes_50m & -0.00 & 0.00 & 0.14 & 40,734 \\
 & pre_treatment_change_in_group_3_crimes_100m & 0.01 & 0.00 & 0.40 & 40,734 \\
 & pre_treatment_change_in_group_3_crimes_150m & 0.01 & 0.00 & 0.57 & 40,734 \\
 & pre_treatment_change_in_group_3_crimes_200m & 0.02 & 0.00 & 0.76 & 40,734 \\
 & pre_treatment_change_in_group_3_crimes_50m & 0.00 & 0.00 & 0.22 & 40,734 \\
 & pre_treatment_change_in_zestimate & 46,593.08 & 37,330.00 & 158,125.24 & 10,160 \\
 & twenty_seventeen_group_0_crimes_100m & 0.77 & 0.00 & 2.89 & 40,734 \\
 & twenty_seventeen_group_0_crimes_150m & 1.59 & 0.00 & 5.61 & 40,734 \\
 & twenty_seventeen_group_0_crimes_200m & 2.70 & 0.00 & 8.82 & 40,734 \\
 & twenty_seventeen_group_0_crimes_50m & 0.22 & 0.00 & 1.11 & 40,734 \\
 & twenty_seventeen_group_1_crimes_100m & 0.06 & 0.00 & 0.34 & 40,734 \\
 & twenty_seventeen_group_1_crimes_150m & 0.12 & 0.00 & 0.59 & 40,734 \\
 & twenty_seventeen_group_1_crimes_200m & 0.21 & 0.00 & 0.83 & 40,734 \\
 & twenty_seventeen_group_1_crimes_50m & 0.02 & 0.00 & 0.16 & 40,734 \\
 & twenty_seventeen_group_2_crimes_100m & 0.03 & 0.00 & 0.37 & 40,734 \\
 & twenty_seventeen_group_2_crimes_150m & 0.06 & 0.00 & 0.56 & 40,734 \\
 & twenty_seventeen_group_2_crimes_200m & 0.11 & 0.00 & 0.86 & 40,734 \\
 & twenty_seventeen_group_2_crimes_50m & 0.01 & 0.00 & 0.26 & 40,734 \\
 & twenty_seventeen_group_3_crimes_100m & 0.07 & 0.00 & 0.36 & 40,734 \\
 & twenty_seventeen_group_3_crimes_150m & 0.14 & 0.00 & 0.63 & 40,734 \\
 & twenty_seventeen_group_3_crimes_200m & 0.25 & 0.00 & 1.02 & 40,734 \\
 & twenty_seventeen_group_3_crimes_50m & 0.02 & 0.00 & 0.16 & 40,734 \\
 & twenty_seventeen_zestimate & 298,898.22 & 235,078.00 & 352,756.46 & 10,176 \\
\cline{1-6}
\multirow[c]{8}{4cm}{\textit{Panel B: Census Tract Characteristics}} & frac_coll_plus2010 & 0.26 & 0.22 & 0.17 & 40,732 \\
 & job_density_2013 & 4,802.71 & 1,355.40 & 17,583.20 & 40,732 \\
 & med_hhinc2016 & 52,695.02 & 47,105.00 & 27,212.35 & 40,732 \\
 & poor_share2010 & 0.20 & 0.16 & 0.15 & 40,732 \\
 & popdensity2010 & 9,151.92 & 5,982.68 & 9,537.28 & 40,732 \\
 & rent_twobed2015 & 1,116.73 & 1,055.00 & 396.69 & 30,546 \\
 & share_white2010 & 0.58 & 0.63 & 0.29 & 40,732 \\
 & traveltime15_2010 & 0.28 & 0.27 & 0.11 & 40,732 \\
\cline{1-6}
\multirow[c]{3}{4cm}{\textit{Panel C: Case Initiation}} & for_cause & 0.12 & 0.00 & 0.33 & 40,734 \\
 & no_cause & 0.11 & 0.00 & 0.31 & 40,734 \\
 & non_payment & 0.75 & 1.00 & 0.43 & 40,734 \\
\cline{1-6}
\multirow[c]{4}{4cm}{\textit{Panel D: Defendant and Plaintiff Characteristics}} & hasAttyD & 0.09 & 0.00 & 0.28 & 40,734 \\
 & hasAttyP & 0.84 & 1.00 & 0.37 & 40,734 \\
 & isEntityD & 0.01 & 0.00 & 0.08 & 40,734 \\
 & isEntityP & 0.70 & 1.00 & 0.46 & 40,734 \\
\cline{1-6}
\multirow[c]{6}{4cm}{\textit{Panel E: Case Resolution}} & case_duration & 57.81 & 21.00 & 78.28 & 39,094 \\
 & defaulted & 0.20 & 0.00 & 0.40 & 40,734 \\
 & dismissed & 0.29 & 0.00 & 0.45 & 40,734 \\
 & heard & 0.06 & 0.00 & 0.23 & 40,734 \\
 & judgment & 1,890.02 & 0.00 & 5,279.08 & 40,734 \\
 & mediated & 0.41 & 0.00 & 0.49 & 40,734 \\
\cline{1-6}
\multirow[c]{34}{4cm}{\textit{Panel F: Post-treatment Outcomes}} & group_0_crimes_100m_1_years_relative_to_treatment & 0.52 & 0.00 & 2.57 & 40,734 \\
 & group_0_crimes_100m_2_years_relative_to_treatment & 0.52 & 0.00 & 2.21 & 35,684 \\
 & group_0_crimes_150m_1_years_relative_to_treatment & 1.08 & 0.00 & 5.19 & 40,734 \\
 & group_0_crimes_150m_2_years_relative_to_treatment & 1.11 & 0.00 & 5.26 & 35,684 \\
 & group_0_crimes_200m_1_years_relative_to_treatment & 1.93 & 0.00 & 8.75 & 40,734 \\
 & group_0_crimes_200m_2_years_relative_to_treatment & 2.01 & 0.00 & 9.13 & 35,684 \\
 & group_0_crimes_50m_1_years_relative_to_treatment & 0.15 & 0.00 & 1.43 & 40,734 \\
 & group_0_crimes_50m_2_years_relative_to_treatment & 0.16 & 0.00 & 0.86 & 35,684 \\
 & group_1_crimes_100m_1_years_relative_to_treatment & 0.04 & 0.00 & 0.28 & 40,734 \\
 & group_1_crimes_100m_2_years_relative_to_treatment & 0.04 & 0.00 & 0.25 & 35,684 \\
 & group_1_crimes_150m_1_years_relative_to_treatment & 0.08 & 0.00 & 0.44 & 40,734 \\
 & group_1_crimes_150m_2_years_relative_to_treatment & 0.08 & 0.00 & 0.42 & 35,684 \\
 & group_1_crimes_200m_1_years_relative_to_treatment & 0.13 & 0.00 & 0.66 & 40,734 \\
 & group_1_crimes_200m_2_years_relative_to_treatment & 0.13 & 0.00 & 0.64 & 35,684 \\
 & group_1_crimes_50m_1_years_relative_to_treatment & 0.01 & 0.00 & 0.16 & 40,734 \\
 & group_1_crimes_50m_2_years_relative_to_treatment & 0.01 & 0.00 & 0.12 & 35,684 \\
 & group_2_crimes_100m_1_years_relative_to_treatment & 0.00 & 0.00 & 0.00 & 40,734 \\
 & group_2_crimes_100m_2_years_relative_to_treatment & 0.00 & 0.00 & 0.00 & 35,684 \\
 & group_2_crimes_150m_1_years_relative_to_treatment & 0.00 & 0.00 & 0.00 & 40,734 \\
 & group_2_crimes_150m_2_years_relative_to_treatment & 0.00 & 0.00 & 0.00 & 35,684 \\
 & group_2_crimes_200m_1_years_relative_to_treatment & 0.00 & 0.00 & 0.00 & 40,734 \\
 & group_2_crimes_200m_2_years_relative_to_treatment & 0.00 & 0.00 & 0.00 & 35,684 \\
 & group_2_crimes_50m_1_years_relative_to_treatment & 0.00 & 0.00 & 0.00 & 40,734 \\
 & group_2_crimes_50m_2_years_relative_to_treatment & 0.00 & 0.00 & 0.00 & 35,684 \\
 & group_3_crimes_100m_1_years_relative_to_treatment & 0.08 & 0.00 & 0.53 & 40,734 \\
 & group_3_crimes_100m_2_years_relative_to_treatment & 0.09 & 0.00 & 0.50 & 35,684 \\
 & group_3_crimes_150m_1_years_relative_to_treatment & 0.16 & 0.00 & 0.79 & 40,734 \\
 & group_3_crimes_150m_2_years_relative_to_treatment & 0.18 & 0.00 & 0.87 & 35,684 \\
 & group_3_crimes_200m_1_years_relative_to_treatment & 0.28 & 0.00 & 1.23 & 40,734 \\
 & group_3_crimes_200m_2_years_relative_to_treatment & 0.33 & 0.00 & 1.50 & 35,684 \\
 & group_3_crimes_50m_1_years_relative_to_treatment & 0.02 & 0.00 & 0.30 & 40,734 \\
 & group_3_crimes_50m_2_years_relative_to_treatment & 0.03 & 0.00 & 0.21 & 35,684 \\
 & zestimate_1_years_relative_to_treatment & 421,590.21 & 347,444.00 & 344,333.24 & 10,443 \\
 & zestimate_2_years_relative_to_treatment & 477,931.84 & 393,800.00 & 457,699.06 & 9,175 \\
\cline{1-6}
\bottomrule
\end{tabular}