### 2023-03-15 writing/presentations/march_15.tex

\documentclass [xcolor=svgnames, t] {beamer} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs, comment} 
\usepackage[absolute, overlay]{textpos} 
\usepackage{pgfpages}
\usepackage[font=footnotesize]{caption}
\useoutertheme{infolines}

\definecolor{brownbrown}{RGB}{56, 28, 0}
\definecolor{brownred}{RGB}{228, 0, 43}

\setbeamercolor{title in head/foot}{bg=brownred, fg=brownbrown}
\setbeamercolor{author in head/foot}{bg=myuniversity}
\setbeamertemplate{page number in head/foot}{}
\usepackage{csquotes}

\usepackage{amsmath}
\usepackage[makeroom]{cancel}

\usepackage{textpos}

\usepackage{tikz}

\usetheme{Madrid}
\definecolor{myuniversity}{RGB}{56, 28, 0}
\usecolortheme[named=myuniversity]{structure}
\usepackage{tikz}

\title[Eviction and Property Values]{Estimating the Effect of Eviction on Assessment Values}

\author[Arjun Shanmugam]{}
\setbeamertemplate{navigation symbols}{}

\date{\today}

\addtobeamertemplate{navigation symbols}{}{%
    \usebeamerfont{footline}%
    \usebeamercolor[fg]{footline}%
    \hspace{1em}%
    \insertframenumber/\inserttotalframenumber
}

\begin{document}
\begin{frame}
\maketitle
\end{frame}

\begin{frame}
\frametitle{Table of Contents}
\tableofcontents
\end{frame}

\section{Introduction}
\begin{frame}{Introduction and Literature}
   \begin{enumerate}
       \item Eviction is a bigger problem in America than in other rich countries
       \begin{enumerate}
           \item Most poor renting families in America
spend over half of their incomes on rent (Desmond, 2017)
       \end{enumerate}
       \item Sociologists have found large effects of evictions using non-experimental methods, but economists 
       \begin{enumerate}
           \item Increased probability of job loss (Desmond, Gershenson, 2016)
           \item Increased probability of depression (Desmond, Kimbro, 2015)
       \end{enumerate} 
       \item However, these estimates are likely confounded
       \begin{enumerate}
           \item ``Ashenfelter dip''
           \item Observable and unobservable confounders
       \end{enumerate}
       \item Only recently have economists began to apply quasi-experimental methods to study eviction
        \begin{enumerate}
            \item Van Dijk et al. find effects on durable consumption, access to credit, and more
        \end{enumerate}
   \end{enumerate}
\end{frame}

\section{Motivation}

\begin{frame}{Motivation}
\begin{enumerate}
    \item Strong negative correlation between the presence of eviction and property values
    \item Several potential channels through which eviction might affect property values
    \begin{enumerate}
        \item Evidence suggests that longer tenancy can increase property values
        \item Or, perhaps landlords are evicting problematic tenants
    \end{enumerate}
    \item Either way—property values determine school district 
    funding, so kids' outcomes could be at stake
    \item Ideally, I would have \emph{market} values instead of assessment values
\end{enumerate}
\end{frame}

\section{Data}
\begin{frame}{Data}
    \begin{enumerate}
        \item Eviction data from MassLandlords, a trade association
        \begin{enumerate}
            \item Scraped around 50,000 eviction case records
            \item Collects case outcome data, including method of resolution, judge name, and more
        \end{enumerate}
        \item Crime data from the Boston Police Department (BPD)
        \begin{enumerate}
            \item Incident-level data for every incident that BPD has responded to since 2015
            \item Coordinates of each incident
        \end{enumerate}
        \item Create 100m radius around every eviction and match to crimes which fall into that radius
    \end{enumerate}
\end{frame}

\begin{frame}{Data}
\begin{table}[H]
    \centering
    \tiny
    \input{output/01_exploratory/tables/summary_statistics.tex}
    \label{tab:my_label}
\end{table}
\end{frame}

\begin{frame}{Data}
\begin{table}[H]
    \centering
    \tiny
    \input{output/01_exploratory/tables/balance_table.tex}
    \label{tab:my_label}
\end{table}
\end{frame}

\section{Empirical Strategy}
\begin{frame}{Empirical Strategy}
    \begin{enumerate}
        \item Most eviction cases in Massachusetts are filed in the Massachusetts Housing Court
        \item Within the Housing Court, cases are assigned to about 15 judges quasi-randomly conditional on city and time
        \item This is the basis for an examiner design study, which uses judge dummies as instruments
    \end{enumerate}
\end{frame}

\begin{frame}{Empirical Strategy}
\begin{table}[H]
    \centering
    \tiny
    \input{output/01_exploratory/tables/placebo_first_stage.tex}
    \label{tab:my_label}
\end{table}    
\end{frame}

\begin{frame}{Preliminary Results}
    \begin{table}[H]
        \centering
        \tiny
        \input{output/01_exploratory/tables/ln_total_val_results_table.tex}
        \label{tab:my_label}
    \end{table}
\end{frame}
\begin{frame}{Preliminary Results}
    \begin{table}[H]
        \centering
        \tiny
        \input{output/01_exploratory/tables/ln_bldg_val_results_table.tex}
        \label{tab:my_label}
    \end{table}
\end{frame}

\begin{frame}{Next Steps}
    \begin{enumerate}
        \item Figure out what to do about my lack of power
        \item Find a systematic way to drop observations in sparse city-years
        \item Speak to an assessor from Massachusetts to understand whether there is even a mechanism through which I might see an effect
        \item Think about whether my instrument is truly valid
    \end{enumerate}
\end{frame}

\end{document}
Close