### 2023-03-23 output/group_3_crimes_100m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & twenty_seventeen_group_3_crimes_100m & 0.45 & 0.05 & 0.08 & -0.00 & 0.95 \\
 & pre_treatment_change_in_group_3_crimes_100m & 0.06 & -0.03 & 0.45 & -0.00 & 0.99 \\
\cline{1-7}
\multirow[c]{3}{3cm}{\textit{Panel B}} & frac_coll_plus2010 & 0.32 & 0.00 & 0.78 & -0.00 & 0.86 \\
 & job_density_2013 & 16,069.14 & 926.91 & 0.52 & -65.36 & 0.96 \\
 & popdensity2010 & 23,086.04 & 1,225.15 & 0.01 & -94.31 & 0.83 \\
\cline{1-7}
\multirow[c]{3}{3cm}{\textit{Panel C}} & for_cause & 0.17 & 0.07 & 0.00 & -0.00 & 0.96 \\
 & no_cause & 0.05 & -0.00 & 0.90 & -0.00 & 0.98 \\
 & non_payment & 0.78 & -0.07 & 0.00 & -0.00 & 0.75 \\
\cline{1-7}
\multirow[c]{3}{3cm}{\textit{Panel D}} & hasAttyP & 0.88 & -0.04 & 0.00 & -0.00 & 0.69 \\
 & isEntityD & 0.01 & -0.01 & 0.01 & -0.00 & 0.98 \\
 & isEntityP & 0.81 & -0.06 & 0.00 & -0.00 & 0.72 \\
\cline{1-7}
\bottomrule
\end{tabular}