### 2023-03-23 writing/presentations/march_15.tex

\documentclass [xcolor=svgnames, t] {beamer} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs, comment} 
\usepackage[absolute, overlay]{textpos} 
\usepackage{pgfpages}
\usepackage[font=footnotesize]{caption}
\useoutertheme{infolines}

\definecolor{brownbrown}{RGB}{56, 28, 0}
\definecolor{brownred}{RGB}{228, 0, 43}

\setbeamercolor{title in head/foot}{bg=brownred, fg=brownbrown}
\setbeamercolor{author in head/foot}{bg=myuniversity}
\setbeamertemplate{page number in head/foot}{}
\usepackage{csquotes}

\usepackage{amsmath}
\usepackage[makeroom]{cancel}

\usepackage{textpos}

\usepackage{tikz}

\usetheme{Madrid}
\definecolor{myuniversity}{RGB}{56, 28, 0}
\usecolortheme[named=myuniversity]{structure}
\usepackage{tikz}

\title[Eviction and Property Values]{Estimating the Effect of Eviction on Assessment Values}

\author[Arjun Shanmugam]{}
\setbeamertemplate{navigation symbols}{}

\date{\today}

\addtobeamertemplate{navigation symbols}{}{%
    \usebeamerfont{footline}%
    \usebeamercolor[fg]{footline}%
    \hspace{1em}%
    \insertframenumber/\inserttotalframenumber
}

\begin{document}
\begin{frame}
\maketitle
\end{frame}

\begin{frame}
\frametitle{Table of Contents}
\tableofcontents
\end{frame}

\section{Introduction}
\begin{frame}{Introduction and Literature}
   \begin{enumerate}
       \item Eviction is a bigger problem in America than in other rich countries
       \begin{enumerate}
           \item Most poor renting families in America
spend over half of their incomes on rent (Desmond, 2017)
       \end{enumerate}
       \item Sociologists have found large effects of evictions using non-experimental methods=
       \item Economists are just beginning to study eviction
       \begin{enumerate}
           \item Increased probability of job loss (Desmond, Gershenson, 2016)
           \item Increased probability of depression (Desmond, Kimbro, 2015)
       \end{enumerate} 
       \item However, these estimates are likely confounded
       \begin{enumerate}
           \item ``Ashenfelter dip''
           \item Observable and unobservable confounders
       \end{enumerate}
       \item Only recently have economists began to apply quasi-experimental methods to study eviction
        \begin{enumerate}
            \item Van Dijk et al. find effects on durable consumption, access to credit, and more
        \end{enumerate}
   \end{enumerate}
\end{frame}

\section{Motivation}

\begin{frame}{Motivation}
\begin{enumerate}
    \item Strong negative correlation between the presence of eviction and crime
    \item Several potential channels through which eviction might affect crime
    \begin{enumerate}
        \item Maybe eviction weakens community bonds that discourage crime
        \item Maybe evicted tenants are more likely to commit crime after receiving notice of eviction
        \item Or, perhaps landlords are evicting problematic tenants
    \end{enumerate}
    \item Either way, understanding effect is important to understanding how eviction is affecting communitiies
    \item Kroeger and La Mattina (2020): Nuisance ordinances increase eviction risk in Ohio
    \item Falcone (2022): Evictions cause crime! But only if nuisance ordinances affect crime only through evictions...
\end{enumerate}
\end{frame}

\section{Data}
\begin{frame}{Data}
    \begin{enumerate}
        \item Eviction data from MassLandlords, a trade association
        \begin{enumerate}
            \item Scraped around 50,000 eviction case records
            \item Collects case outcome data, including method of resolution, judge name, and more
        \end{enumerate}
        \item Crime data from the Boston Police Department (BPD)
        \begin{enumerate}
            \item Incident-level data for every incident that BPD has responded to since 2015
            \item Coordinates of each incident
        \end{enumerate}
        \item Create 100m radius around every eviction and match to crimes which fall into that radius
    \end{enumerate}
\end{frame}

\begin{frame}{Data}
\begin{table}[H]
    \centering
    \tiny
    \input{output/01_exploratory/tables/summary_statistics.tex}
    \label{tab:my_label}
\end{table}
\end{frame}

\begin{frame}{Data}
\begin{table}[H]
    \centering
    \tiny
    \input{output/01_exploratory/tables/balance_table.tex}
    \label{tab:my_label}
\end{table}
\end{frame}

\section{Empirical Strategy}
\begin{frame}{Empirical Strategy}
    \begin{enumerate}
        \item Most eviction cases in Massachusetts are filed in the Massachusetts Housing Court
        \item Within the Housing Court, cases are assigned to about 15 judges quasi-randomly conditional on city and time
        \item This is the basis for an examiner design study, which uses judge dummies as instruments
    \end{enumerate}
\end{frame}

\begin{frame}{Empirical Strategy}
\begin{table}[H]
    \centering
    \tiny
    \input{output/01_exploratory/tables/placebo_first_stage.tex}
    \label{tab:my_label}
\end{table}    
\end{frame}

\begin{frame}{Preliminary Results}
    \begin{table}[H]
        \centering
        \tiny
        \input{output/01_exploratory/tables/ln_total_val_results_table.tex}
        \label{tab:my_label}
    \end{table}
\end{frame}
\begin{frame}{Preliminary Results}
    \begin{table}[H]
        \centering
        \tiny
        \input{output/01_exploratory/tables/ln_bldg_val_results_table.tex}
        \label{tab:my_label}
    \end{table}
\end{frame}

\begin{frame}{Next Steps}
    \begin{enumerate}
        \item Figure out what to do about my lack of power
        \item Find a systematic way to drop observations in sparse city-years
        \item Speak to an assessor from Massachusetts to understand whether there is even a mechanism through which I might see an effect
        \item Think about whether my instrument is truly valid
    \end{enumerate}
\end{frame}

\end{document}
Close