### 2023-03-31 output/group_0_crimes_100m/tables/balance_table.tex
biweekly analysis

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & total_twenty_seventeen_group_2_crimes_100m & 7.20 & 0.54 & 0.04 & -0.05 & 0.83 \\
 & pre_treatment_change_in_group_2_crimes_100m & 0.03 & -0.03 & 0.13 & -0.00 & 0.95 \\
\cline{1-7}
\multirow[c]{4}{3cm}{\textit{Panel B}} & frac_coll_plus2010 & 0.32 & 0.00 & 0.78 & -0.00 & 0.68 \\
 & med_hhinc2016 & 48,035.65 & 1,312.56 & 0.12 & -429.96 & 0.60 \\
 & poor_share2010 & 0.27 & -0.00 & 0.53 & -0.00 & 0.57 \\
 & popdensity2010 & 23,086.04 & 1,225.15 & 0.01 & -185.35 & 0.65 \\
\cline{1-7}
\multirow[c]{2}{3cm}{\textit{Panel D}} & isEntityD & 0.01 & -0.01 & 0.01 & -0.00 & 0.90 \\
 & isEntityP & 0.81 & -0.06 & 0.00 & -0.01 & 0.30 \\
\cline{1-7}
\bottomrule
\end{tabular}