### 2023-03-31 output/summary_statistics/tables/summary_statistics.tex

\begin{tabular}{llcccc}
\toprule
 &  & Mean & Median & S.D. & N \\
Panel & Variable &  &  &  &  \\
\midrule
\multirow[c]{6}{4cm}{\textit{Panel A: Pre-treatment Outcomes}} & pre_treatment_change_in_group_0_crimes_100m & -0.30 & 0.00 & 2.38 & 6,060 \\
 & pre_treatment_change_in_group_1_crimes_100m & 0.02 & 0.00 & 0.66 & 6,060 \\
 & pre_treatment_change_in_group_2_crimes_100m & 0.03 & 0.00 & 0.68 & 6,060 \\
 & twenty_seventeen_group_0_crimes_100m & 2.32 & 2.00 & 2.83 & 6,060 \\
 & twenty_seventeen_group_1_crimes_100m & 0.17 & 0.00 & 0.47 & 6,060 \\
 & twenty_seventeen_group_2_crimes_100m & 0.19 & 0.00 & 0.47 & 6,060 \\
\cline{1-6}
\multirow[c]{5}{4cm}{\textit{Panel B: Census Tract Characteristics}} & frac_coll_plus2010 & 0.31 & 0.24 & 0.22 & 6,060 \\
 & job_density_2013 & 14,806.08 & 2,425.15 & 40,595.37 & 6,060 \\
 & med_hhinc2016 & 47,188.87 & 41,146.00 & 24,703.65 & 6,060 \\
 & poor_share2010 & 0.28 & 0.25 & 0.14 & 6,060 \\
 & popdensity2010 & 22,704.52 & 20,644.44 & 13,091.45 & 6,060 \\
\cline{1-6}
\multirow[c]{3}{4cm}{\textit{Panel C: Case Initiation}} & for_cause & 0.17 & 0.00 & 0.37 & 6,060 \\
 & no_cause & 0.05 & 0.00 & 0.22 & 6,060 \\
 & non_payment & 0.78 & 1.00 & 0.42 & 6,060 \\
\cline{1-6}
\multirow[c]{4}{4cm}{\textit{Panel D: Defendant and Plaintiff Characteristics}} & hasAttyD & 0.08 & 0.00 & 0.27 & 6,060 \\
 & hasAttyP & 0.88 & 1.00 & 0.32 & 6,060 \\
 & isEntityD & 0.02 & 0.00 & 0.14 & 6,060 \\
 & isEntityP & 0.81 & 1.00 & 0.39 & 6,060 \\
\cline{1-6}
\multirow[c]{6}{4cm}{\textit{Panel E: Case Resolution}} & case_duration & 58.17 & 19.00 & 88.96 & 5,500 \\
 & defaulted & 0.19 & 0.00 & 0.39 & 6,060 \\
 & dismissed & 0.27 & 0.00 & 0.44 & 6,060 \\
 & heard & 0.03 & 0.00 & 0.17 & 6,060 \\
 & judgment & 1,938.57 & 0.00 & 5,337.87 & 6,060 \\
 & mediated & 0.40 & 0.00 & 0.49 & 6,060 \\
\cline{1-6}
\multirow[c]{6}{4cm}{\textit{Panel F: Post-treatment Outcomes}} & group_0_crimes_100m_1_years_relative_to_treatment & 1.46 & 1.00 & 3.05 & 3,039 \\
 & group_0_crimes_100m_2_years_relative_to_treatment & 1.48 & 1.00 & 2.31 & 2,830 \\
 & group_1_crimes_100m_1_years_relative_to_treatment & 0.14 & 0.00 & 0.45 & 3,039 \\
 & group_1_crimes_100m_2_years_relative_to_treatment & 0.10 & 0.00 & 0.36 & 2,830 \\
 & group_2_crimes_100m_1_years_relative_to_treatment & 0.20 & 0.00 & 0.55 & 3,039 \\
 & group_2_crimes_100m_2_years_relative_to_treatment & 0.22 & 0.00 & 0.63 & 2,830 \\
\cline{1-6}
\bottomrule
\end{tabular}