### 2023-04-02 output/group_0_crimes_100m/tables/balance_table.tex
edited merge script

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & total_twenty_seventeen_group_4_crimes_100m & 1.67 & -0.05 & 0.74 & -0.07 & 0.62 \\
 & pre_treatment_change_in_group_4_crimes_100m & -0.02 & -0.00 & 0.87 & 0.00 & 1.00 \\
\cline{1-7}
\multirow[c]{3}{3cm}{\textit{Panel B}} & frac_coll_plus2010 & 0.32 & 0.00 & 0.78 & -0.01 & 0.11 \\
 & job_density_2013 & 16,069.14 & 926.91 & 0.52 & -234.26 & 0.87 \\
 & med_hhinc2016 & 48,035.65 & 1,312.56 & 0.12 & -1,185.57 & 0.00 \\
\cline{1-7}
\textit{Panel D} & isEntityP & 0.81 & -0.06 & 0.00 & -0.05 & 0.00 \\
\cline{1-7}
\bottomrule
\end{tabular}