### 2023-04-03 output/group_0_crimes_100m/tables/balance_table.tex
rann aalysis

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & total_twenty_seventeen_group_0_crimes_100m & 69.43 & 5.31 & 0.03 & -0.53 & 0.82 \\
 & pre_treatment_change_in_group_0_crimes_100m & -0.23 & 0.00 & 0.96 & 0.00 & 0.98 \\
\cline{1-7}
\multirow[c]{5}{3cm}{\textit{Panel B}} & frac_coll_plus2010 & 0.33 & 0.01 & 0.22 & -0.00 & 0.75 \\
 & job_density_2013 & 17,324.75 & 2,509.70 & 0.10 & -133.88 & 0.93 \\
 & med_hhinc2016 & 48,534.18 & 1,788.07 & 0.05 & -373.19 & 0.67 \\
 & poor_share2010 & 0.28 & -0.00 & 0.96 & -0.00 & 0.69 \\
 & popdensity2010 & 23,285.59 & 1,452.05 & 0.00 & -179.26 & 0.69 \\
\cline{1-7}
\textit{Panel C} & non_payment & 0.77 & -0.08 & 0.00 & -0.01 & 0.61 \\
\cline{1-7}
\multirow[c]{3}{3cm}{\textit{Panel D}} & hasAttyP & 0.88 & -0.04 & 0.00 & -0.01 & 0.52 \\
 & isEntityD & 0.02 & -0.01 & 0.06 & -0.00 & 0.98 \\
 & isEntityP & 0.81 & -0.07 & 0.00 & -0.01 & 0.59 \\
\cline{1-7}
\bottomrule
\end{tabular}