### 2023-04-03 output/group_1_crimes_50m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & total_twenty_seventeen_group_1_crimes_50m & 1.72 & 0.20 & 0.05 & 0.02 & 0.86 \\
 & pre_treatment_change_in_group_1_crimes_50m & 0.01 & -0.00 & 0.84 & 0.00 & 0.99 \\
\cline{1-7}
\multirow[c]{4}{3cm}{\textit{Panel B}} & job_density_2013 & 17,324.75 & 2,509.70 & 0.10 & 149.36 & 0.92 \\
 & med_hhinc2016 & 48,534.18 & 1,788.07 & 0.05 & 640.38 & 0.47 \\
 & poor_share2010 & 0.28 & -0.00 & 0.96 & 0.00 & 0.43 \\
 & popdensity2010 & 23,285.59 & 1,452.05 & 0.00 & 282.08 & 0.50 \\
\cline{1-7}
\textit{Panel D} & hasAttyD & 0.07 & -0.03 & 0.00 & 0.00 & 0.75 \\
\cline{1-7}
\bottomrule
\end{tabular}