### 2023-04-03 output/group_2_crimes_50m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & total_twenty_seventeen_group_2_crimes_50m & 0.79 & 0.16 & 0.15 & -0.02 & 0.77 \\
 & pre_treatment_change_in_group_2_crimes_50m & -0.01 & 0.00 & 0.67 & -0.00 & 0.91 \\
\cline{1-7}
\multirow[c]{3}{3cm}{\textit{Panel B}} & frac_coll_plus2010 & 0.33 & 0.01 & 0.22 & 0.00 & 0.62 \\
 & job_density_2013 & 17,324.75 & 2,509.70 & 0.10 & -279.43 & 0.79 \\
 & med_hhinc2016 & 48,534.18 & 1,788.07 & 0.05 & 472.04 & 0.52 \\
\cline{1-7}
\bottomrule
\end{tabular}