### 2023-04-03 output/group_4_crimes_100m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\textit{Panel A} & total_twenty_seventeen_group_4_crimes_100m & 0.06 & 0.02 & 0.01 & -0.00 & 0.88 \\
\cline{1-7}
\bottomrule
\end{tabular}