### 2023-04-06 output/group_2_crimes_1000m/tables/pre_treatment_covariate_tests.tex

\begin{tabular}{llcc}
\toprule
 &  & \multicolumn{2}{c}{\textit{Dependent Variable}} \\
\cline{3-4}
\\
 &  &  &  \\
 & \emph{Independent Variable} &  &  \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A: Pre-treatment Outcomes}} & total_twenty_seventeen_group_2_crimes_1000m & 0.00 & 0.18 \\
 & pre_treatment_change_in_group_2_crimes_1000m & 0.00 & 0.49 \\
\cline{1-4}
\multirow[c]{5}{3cm}{\textit{Panel B: Census Tract Characteristics}} & frac_coll_plus2010 & 0.00 & 0.22 \\
 & job_density_2013 & 0.00 & 0.10 \\
 & med_hhinc2016 & 0.00 & 0.05 \\
 & poor_share2010 & 0.00 & 0.96 \\
 & popdensity2010 & 0.00 & 0.00 \\
\cline{1-4}
\multirow[c]{3}{3cm}{\textit{Panel C: Case Initiation}} & for_cause & 0.37 & 0.00 \\
 & no_cause & 0.16 & 0.95 \\
 & non_payment & 0.11 & 0.00 \\
\cline{1-4}
\multirow[c]{4}{3cm}{\textit{Panel D: Defendant and Plaintiff Characteristics}} & hasAttyD & 0.66 & 0.00 \\
 & hasAttyP & 0.14 & 0.00 \\
 & isEntityD & 0.00 & 0.06 \\
 & isEntityP & 0.54 & 0.00 \\
\cline{1-4}
\bottomrule
\end{tabular}