### 2023-04-06 output/group_2_crimes_100m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & total_twenty_seventeen_group_2_crimes_100m & 0.96 & 0.08 & 0.64 & -0.00 & 0.99 \\
 & pre_treatment_change_in_group_2_crimes_100m & -0.02 & -0.01 & 0.23 & 0.00 & 0.73 \\
\cline{1-7}
\multirow[c]{3}{3cm}{\textit{Panel B}} & frac_coll_plus2010 & 0.33 & 0.01 & 0.22 & 0.00 & 0.64 \\
 & med_hhinc2016 & 48,534.18 & 1,788.07 & 0.05 & 435.35 & 0.52 \\
 & poor_share2010 & 0.28 & -0.00 & 0.96 & 0.01 & 0.27 \\
\cline{1-7}
\bottomrule
\end{tabular}