### 2023-04-06 output/group_3_crimes_100m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & total_twenty_seventeen_group_3_crimes_100m & 1.09 & 0.22 & 0.04 & 0.05 & 0.63 \\
 & pre_treatment_change_in_group_3_crimes_100m & -0.05 & -0.01 & 0.54 & -0.00 & 0.85 \\
\cline{1-7}
\multirow[c]{2}{3cm}{\textit{Panel B}} & job_density_2013 & 17,324.75 & 6,144.92 & 0.01 & 2,223.13 & 0.36 \\
 & poor_share2010 & 0.28 & 0.02 & 0.02 & -0.01 & 0.31 \\
\cline{1-7}
\textit{Panel D} & isEntityP & 0.81 & 0.13 & 0.00 & 0.02 & 0.19 \\
\cline{1-7}
\bottomrule
\end{tabular}