### 2023-04-06 output/group_6_crimes_1000m/tables/balance_table.tex
Updated constants

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & total_twenty_seventeen_group_4_crimes_1000m & 3.40 & 0.30 & 0.00 & -0.01 & 0.93 \\
 & pre_treatment_change_in_group_4_crimes_1000m & 0.17 & -0.05 & 0.03 & -0.00 & 0.98 \\
\cline{1-7}
\multirow[c]{5}{3cm}{\textit{Panel B}} & frac_coll_plus2010 & 0.33 & 0.01 & 0.22 & -0.00 & 0.92 \\
 & job_density_2013 & 17,324.75 & 2,509.70 & 0.10 & -34.03 & 0.98 \\
 & med_hhinc2016 & 48,534.18 & 1,788.07 & 0.05 & -110.46 & 0.90 \\
 & poor_share2010 & 0.28 & -0.00 & 0.96 & -0.00 & 0.90 \\
 & popdensity2010 & 23,285.59 & 1,452.05 & 0.00 & -51.28 & 0.91 \\
\cline{1-7}
\multirow[c]{3}{3cm}{\textit{Panel D}} & hasAttyP & 0.88 & -0.04 & 0.00 & -0.00 & 0.83 \\
 & isEntityD & 0.02 & -0.01 & 0.06 & -0.00 & 0.99 \\
 & isEntityP & 0.81 & -0.07 & 0.00 & -0.00 & 0.85 \\
\cline{1-7}
\bottomrule
\end{tabular}