### 2023-04-08 output/group_2_crimes_1000m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & total_twenty_seventeen_group_2_crimes_1000m & 59.19 & 3.58 & 0.18 & 0.65 & 0.80 \\
 & pre_treatment_change_in_group_2_crimes_1000m & -0.67 & -0.05 & 0.50 & -0.01 & 0.93 \\
\cline{1-7}
\multirow[c]{5}{3cm}{\textit{Panel B}} & frac_coll_plus2010 & 0.33 & 0.01 & 0.22 & 0.00 & 0.60 \\
 & job_density_2013 & 17,324.75 & 2,509.70 & 0.10 & 113.70 & 0.94 \\
 & med_hhinc2016 & 48,534.18 & 1,788.07 & 0.05 & 593.73 & 0.49 \\
 & poor_share2010 & 0.28 & -0.00 & 0.96 & 0.00 & 0.44 \\
 & popdensity2010 & 23,285.59 & 1,452.05 & 0.00 & 253.67 & 0.52 \\
\cline{1-7}
\textit{Panel D} & isEntityD & 0.02 & -0.01 & 0.06 & 0.00 & 0.87 \\
\cline{1-7}
\bottomrule
\end{tabular}