### 2023-04-08 output/group_3_crimes_100m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\textit{Panel A} & total_twenty_seventeen_group_3_crimes_100m & 7.20 & 0.82 & 0.04 & -0.03 & 0.94 \\
\cline{1-7}
\multirow[c]{4}{3cm}{\textit{Panel B}} & frac_coll_plus2010 & 0.33 & 0.02 & 0.09 & -0.01 & 0.52 \\
 & job_density_2013 & 17,324.75 & 6,144.92 & 0.01 & 1,756.80 & 0.47 \\
 & med_hhinc2016 & 48,534.18 & 2,249.48 & 0.13 & -1,615.62 & 0.22 \\
 & popdensity2010 & 23,285.59 & 1,707.93 & 0.03 & -500.51 & 0.47 \\
\cline{1-7}
\textit{Panel D} & isEntityD & 0.02 & -0.04 & 0.00 & -0.02 & 0.01 \\
\cline{1-7}
\bottomrule
\end{tabular}