### 2023-04-08 output/group_3_crimes_50m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\textit{Panel A} & total_twenty_seventeen_group_3_crimes_50m & 2.07 & 0.22 & 0.07 & -0.01 & 0.31 \\
\cline{1-7}
\bottomrule
\end{tabular}