### 2023-04-08 output/group_4_crimes_50m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & total_twenty_seventeen_group_4_crimes_50m & 1.77 & -0.10 & 0.54 & -0.05 & 0.77 \\
 & pre_treatment_change_in_group_4_crimes_50m & 0.01 & 0.01 & 0.48 & 0.00 & 1.00 \\
\cline{1-7}
\multirow[c]{3}{3cm}{\textit{Panel B}} & frac_coll_plus2010 & 0.33 & 0.01 & 0.22 & -0.01 & 0.26 \\
 & job_density_2013 & 17,324.75 & 2,509.70 & 0.10 & -391.41 & 0.79 \\
 & popdensity2010 & 23,285.59 & 1,452.05 & 0.00 & -578.94 & 0.11 \\
\cline{1-7}
\multirow[c]{2}{3cm}{\textit{Panel D}} & isEntityD & 0.02 & -0.01 & 0.06 & -0.00 & 0.89 \\
 & isEntityP & 0.81 & -0.07 & 0.00 & -0.02 & 0.04 \\
\cline{1-7}
\bottomrule
\end{tabular}