### 2023-04-08 output/group_5_crimes_100m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & total_twenty_seventeen_group_5_crimes_100m & 3.87 & 0.18 & 0.24 & -0.05 & 0.75 \\
 & pre_treatment_change_in_group_5_crimes_100m & 0.01 & -0.01 & 0.52 & -0.00 & 0.98 \\
\cline{1-7}
\multirow[c]{3}{3cm}{\textit{Panel B}} & frac_coll_plus2010 & 0.33 & 0.01 & 0.22 & -0.00 & 0.60 \\
 & job_density_2013 & 17,324.75 & 2,509.70 & 0.10 & -159.16 & 0.91 \\
 & popdensity2010 & 23,285.59 & 1,452.05 & 0.00 & -261.65 & 0.45 \\
\cline{1-7}
\textit{Panel D} & hasAttyP & 0.88 & -0.04 & 0.00 & -0.01 & 0.21 \\
\cline{1-7}
\bottomrule
\end{tabular}