### 2023-04-08 output/group_5_crimes_50m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & total_twenty_seventeen_group_5_crimes_50m & 1.13 & 0.03 & 0.69 & 0.04 & 0.55 \\
 & pre_treatment_change_in_group_5_crimes_50m & 0.01 & -0.00 & 0.72 & 0.00 & 0.71 \\
\cline{1-7}
\textit{Panel B} & popdensity2010 & 23,285.59 & 1,452.05 & 0.00 & 88.93 & 0.51 \\
\cline{1-7}
\bottomrule
\end{tabular}