### 2023-04-08 output/group_6_crimes_50m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel B}} & job_density_2013 & 17,324.75 & 2,509.70 & 0.10 & 345.62 & 0.82 \\
 & med_hhinc2016 & 48,534.18 & 1,788.07 & 0.05 & 1,321.57 & 0.15 \\
\cline{1-7}
\textit{Panel C} & for_cause & 0.17 & 0.08 & 0.00 & 0.00 & 0.98 \\
\cline{1-7}
\bottomrule
\end{tabular}