### 2023-04-10 output/group_0_crimes_1000m/tables/balance_table.tex
Easy switching between monthly and weekly analysis

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & Total Group 0 Incidents, 2017 & 3,981.99 & 163.75 & 0.04 & -99.78 & 0.18 \\
 & $\Delta$ Group 0 Incidents, 2017-2019 & -490.95 & -20.32 & 0.20 & 12.30 & 0.43 \\
\cline{1-7}
\multirow[c]{4}{3cm}{\textit{Panel B}} & Bachelor's degree, 2010 & 0.33 & 0.01 & 0.22 & -0.01 & 0.28 \\
 & Job density, 2013 & 17,324.75 & 2,509.70 & 0.10 & -398.45 & 0.79 \\
 & Poverty rate, 2010 & 0.28 & -0.00 & 0.96 & -0.01 & 0.17 \\
 & Population density, 2010 & 23,285.59 & 1,452.05 & 0.00 & -573.69 & 0.16 \\
\cline{1-7}
\multirow[c]{2}{3cm}{\textit{Panel C}} & Filing without cause & 0.05 & 0.00 & 0.95 & -0.00 & 0.87 \\
 & Filing for nonpayment & 0.77 & -0.08 & 0.00 & -0.02 & 0.08 \\
\cline{1-7}
\multirow[c]{2}{3cm}{\textit{Panel D}} & Defendant has attorney & 0.07 & -0.03 & 0.00 & -0.00 & 0.80 \\
 & Defendant is entity & 0.02 & -0.01 & 0.06 & -0.00 & 0.90 \\
\cline{1-7}
\bottomrule
\end{tabular}