### 2023-04-10 output/group_2_crimes_1000m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & Total Group 2 Incidents, 2017 & 58.13 & 3.46 & 0.19 & 0.12 & 0.96 \\
 & $\Delta$ Group 2 Incidents, 2017-2019 & -10.95 & -0.05 & 0.92 & -0.03 & 0.96 \\
\cline{1-7}
\multirow[c]{5}{3cm}{\textit{Panel B}} & Bachelor's degree, 2010 & 0.33 & 0.01 & 0.22 & 0.00 & 0.93 \\
 & Job density, 2013 & 17,324.75 & 2,509.70 & 0.10 & 29.56 & 0.98 \\
 & Median household income, 2016 & 48,534.18 & 1,788.07 & 0.05 & 104.87 & 0.91 \\
 & Poverty rate, 2010 & 0.28 & -0.00 & 0.96 & 0.00 & 0.90 \\
 & Population density, 2010 & 23,285.59 & 1,452.05 & 0.00 & 47.81 & 0.91 \\
\cline{1-7}
\multirow[c]{2}{3cm}{\textit{Panel D}} & Defendant is entity & 0.02 & -0.01 & 0.06 & 0.00 & 0.99 \\
 & Plaintiff is entity & 0.81 & -0.07 & 0.00 & 0.00 & 0.82 \\
\cline{1-7}
\bottomrule
\end{tabular}