### 2023-04-10 output/group_3_crimes_1000m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\textit{Panel A} & Total Group 3 Incidents, 2017 & 410.15 & 15.31 & 0.04 & -1.86 & 0.79 \\
\cline{1-7}
\multirow[c]{4}{3cm}{\textit{Panel B}} & Bachelor's degree, 2010 & 0.33 & 0.01 & 0.22 & -0.00 & 0.85 \\
 & Job density, 2013 & 17,324.75 & 2,509.70 & 0.10 & -66.40 & 0.96 \\
 & Median household income, 2016 & 48,534.18 & 1,788.07 & 0.05 & -220.39 & 0.80 \\
 & Population density, 2010 & 23,285.59 & 1,452.05 & 0.00 & -101.84 & 0.81 \\
\cline{1-7}
\multirow[c]{2}{3cm}{\textit{Panel D}} & Plaintiff has attorney & 0.88 & -0.04 & 0.00 & -0.00 & 0.65 \\
 & Plaintiff is entity & 0.81 & -0.07 & 0.00 & -0.00 & 0.65 \\
\cline{1-7}
\bottomrule
\end{tabular}