### 2023-04-10 output/summary_statistics/tables/summary_statistics.tex

\begin{tabular}{llcccc}
\toprule
 &  & Mean & S.D. & N \\
Panel & Variable &  &  &  \\
\midrule
\multirow[c]{2}{4cm}{\textit{Panel A: Pre-treatment Outcomes}} & $\Delta$ Group 0 Incidents, 2017-2019 & -532.85 & 465.54 & 6,060 \\
 & Total Group 0 Incidents, 2017 & 4,068.25 & 2,258.29 & 6,060 \\
\cline{1-5}
\multirow[c]{5}{4cm}{\textit{Panel B: Census Tract Characteristics}} & Bachelor's degree, 2010 & 0.31 & 0.22 & 6,060 \\
 & Job density, 2013 & 14,806.08 & 40,595.37 & 6,060 \\
 & Median household income, 2016 & 47,188.87 & 24,703.65 & 6,060 \\
 & Poverty rate, 2010 & 0.28 & 0.14 & 6,060 \\
 & Population density, 2010 & 22,704.52 & 13,091.45 & 6,060 \\
\cline{1-5}
\multirow[c]{3}{4cm}{\textit{Panel C: Case Initiation}} & Filing for cause & 0.17 & 0.37 & 6,060 \\
 & Filing without cause & 0.05 & 0.22 & 6,060 \\
 & Filing for nonpayment & 0.78 & 0.42 & 6,060 \\
\cline{1-5}
\multirow[c]{4}{4cm}{\textit{Panel D: Defendant and Plaintiff Characteristics}} & Defendant has attorney & 0.08 & 0.27 & 6,060 \\
 & Plaintiff has attorney & 0.88 & 0.32 & 6,060 \\
 & Defendant is entity & 0.02 & 0.14 & 6,060 \\
 & Plaintiff is entity & 0.81 & 0.39 & 6,060 \\
\cline{1-5}
\multirow[c]{6}{4cm}{\textit{Panel E: Case Resolution}} & Case duration & 58.17 & 88.96 & 5,500 \\
 & Judgment by default & 0.19 & 0.39 & 6,060 \\
 & Case dismissed & 0.27 & 0.44 & 6,060 \\
 & Case heard & 0.03 & 0.17 & 6,060 \\
 & Money judgment & 1,938.57 & 5,337.87 & 6,060 \\
 & Case mediated & 0.40 & 0.49 & 6,060 \\
\cline{1-5}
\bottomrule
\end{tabular}