### 2023-04-10 output/group_1_crimes_500m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A}} & Total Group 1 Incidents, 2017 & 95.23 & 5.45 & 0.02 & 1.12 & 0.59 \\
 & $\Delta$ Group 1 Incidents, 2017-2019 & 2.34 & 1.50 & 0.09 & -0.05 & 0.96 \\
\cline{1-7}
\multirow[c]{5}{3cm}{\textit{Panel B}} & Bachelor's degree, 2010 & 0.33 & 0.01 & 0.22 & 0.00 & 0.58 \\
 & Job density, 2013 & 17,324.75 & 2,509.70 & 0.10 & 122.05 & 0.93 \\
 & Median household income, 2016 & 48,534.18 & 1,788.07 & 0.05 & 623.35 & 0.47 \\
 & Poverty rate, 2010 & 0.28 & -0.00 & 0.96 & 0.00 & 0.42 \\
 & Population density, 2010 & 23,285.59 & 1,452.05 & 0.00 & 267.17 & 0.50 \\
\cline{1-7}
\textit{Panel D} & Defendant is entity & 0.02 & -0.01 & 0.06 & 0.00 & 0.87 \\
\cline{1-7}
\bottomrule
\end{tabular}