### 2023-04-12 output/group_1_crimes_500m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{.75cm}{\textit{Panel A}} & Total Group 1 Incidents, 2017 & 2.98 & 0.26 & 0.10 & 0.00 & 0.98 \\
 & $\Delta$ Group 1 Incidents, 2017-2019 & 12.72 & 0.07 & 0.92 & -0.18 & 0.79 \\
\cline{1-7}
\multirow[c]{5}{.75cm}{\textit{Panel B}} & Bachelor's degree, 2010 & 0.32 & 0.02 & 0.03 & -0.00 & 0.98 \\
 & Job density, 2013 & 17,879.13 & 2,380.87 & 0.29 & 184.04 & 0.93 \\
 & Median household income, 2016 & 47,493.96 & 3,485.70 & 0.01 & -54.60 & 0.94 \\
 & Poverty rate, 2010 & 0.28 & -0.02 & 0.01 & -0.01 & 0.22 \\
 & Population density, 2010 & 23,692.02 & 403.43 & 0.57 & -282.90 & 0.68 \\
\cline{1-7}
\bottomrule
\end{tabular}