### 2023-04-12 output/summary_statistics/tables/summary_statistics.tex

\begin{tabular}{llcccc}
\toprule
 &  & Mean & S.D. & N \\
Panel & Variable &  &  &  \\
\midrule
\multirow[c]{2}{4cm}{\textit{Panel A: Pre-treatment Outcomes}} & Change in Crime Incidents, 2017-2019 & -164.98 & 222.44 & 3,171 \\
 & Total Crime Incidents, 2017 & 1,230.16 & 697.53 & 3,171 \\
\cline{1-5}
\multirow[c]{5}{4cm}{\textit{Panel B: Census Tract Characteristics}} & Bachelor's degree, 2010 & 0.30 & 0.21 & 3,171 \\
 & Job density, 2013 & 15,603.89 & 41,982.90 & 3,171 \\
 & Median household income, 2016 & 46,085.28 & 24,946.22 & 3,171 \\
 & Poverty rate, 2010 & 0.29 & 0.15 & 3,171 \\
 & Population density, 2010 & 23,385.01 & 13,324.90 & 3,171 \\
\cline{1-5}
\multirow[c]{3}{4cm}{\textit{Panel C: Case Initiation}} & Filing for cause & 0.16 & 0.36 & 3,171 \\
 & Filing without cause & 0.05 & 0.23 & 3,171 \\
 & Filing for nonpayment & 0.79 & 0.41 & 3,171 \\
\cline{1-5}
\multirow[c]{4}{4cm}{\textit{Panel D: Defendant and Plaintiff Characteristics}} & Defendant has attorney & 0.04 & 0.20 & 3,171 \\
 & Plaintiff has attorney & 0.87 & 0.34 & 3,171 \\
 & Defendant is entity & 0.02 & 0.13 & 3,171 \\
 & Plaintiff is entity & 0.82 & 0.39 & 3,171 \\
\cline{1-5}
\multirow[c]{6}{4cm}{\textit{Panel E: Case Resolution}} & Case duration & 22.33 & 23.38 & 3,039 \\
 & Judgment by default & 0.25 & 0.43 & 3,171 \\
 & Case dismissed & 0.20 & 0.40 & 3,171 \\
 & Case heard & 0.02 & 0.14 & 3,171 \\
 & Money judgment & 1,806.43 & 4,135.66 & 3,171 \\
 & Case mediated & 0.48 & 0.50 & 3,171 \\
\cline{1-5}
\bottomrule
\end{tabular}