### 2023-04-12 output/summary_statistics/tables/summary_statistics.tex

\begin{tabular}{llcccc}
\toprule
 &  & Mean & S.D. & N \\
Panel & Variable &  &  &  \\
\midrule
\multirow[c]{2}{4cm}{\textit{Panel A: Pre-treatment Outcomes}} & Change in Crime Incidents, 2017-2019 & -152.00 & 217.38 & 1,632 \\
 & Total Crime Incidents, 2017 & 1,162.73 & 714.55 & 1,632 \\
\cline{1-5}
\multirow[c]{5}{4cm}{\textit{Panel B: Census Tract Characteristics}} & Bachelor's degree, 2010 & 0.31 & 0.22 & 1,632 \\
 & Job density, 2013 & 16,511.95 & 43,184.21 & 1,632 \\
 & Median household income, 2016 & 46,399.34 & 25,059.82 & 1,632 \\
 & Poverty rate, 2010 & 0.29 & 0.15 & 1,632 \\
 & Population density, 2010 & 23,373.29 & 14,260.04 & 1,632 \\
\cline{1-5}
\multirow[c]{3}{4cm}{\textit{Panel C: Case Initiation}} & Filing for cause & 0.11 & 0.32 & 1,632 \\
 & Filing without cause & 0.04 & 0.20 & 1,632 \\
 & Filing for nonpayment & 0.76 & 0.43 & 1,632 \\
\cline{1-5}
\multirow[c]{4}{4cm}{\textit{Panel D: Defendant and Plaintiff Characteristics}} & Defendant has attorney & 0.03 & 0.18 & 1,632 \\
 & Plaintiff has attorney & 0.91 & 0.29 & 1,632 \\
 & Defendant is entity & 0.02 & 0.15 & 1,632 \\
 & Plaintiff is entity & 0.84 & 0.36 & 1,632 \\
\cline{1-5}
\multirow[c]{10}{4cm}{\textit{Panel E: Case Resolution}} & Case duration & 19.82 & 24.51 & 1,538 \\
 & Judgment by default & 0.48 & 0.50 & 1,632 \\
 & Case dismissed & 0.41 & 0.49 & 1,632 \\
 & Case heard & 0.05 & 0.21 & 1,632 \\
 & Execution issued & 0.30 & 0.46 & 1,632 \\
 & Money judgment & 1,456.98 & 3,445.07 & 1,632 \\
 & Execution levied & 0.06 & 0.23 & 1,632 \\
 & Case mediated & 0.00 & 0.00 & 1,632 \\
 & Execution requested & 0.02 & 0.15 & 1,632 \\
 & Execution status unknown & 0.00 & 0.00 & 1,632 \\
\cline{1-5}
\bottomrule
\end{tabular}