### 2023-04-12 output/group_0_crimes_500m/tables/pre_treatment_covariate_tests.tex
generated figures

\begin{tabular}{llcc}
\toprule
 &  & \multicolumn{2}{c}{\textit{Dependent Variable}} \\
\cline{3-4}
\\
 &  &  &  \\
 & \emph{Independent Variable} &  &  \\
\midrule
\multirow[c]{2}{3cm}{\textit{Panel A: Pre-treatment Outcomes}} & Total Crime Incidents, 2017 & 0.00 & 0.47 \\
 & Change in Crime Incidents, 2017-2019 & 0.00 & 0.46 \\
\cline{1-4}
\multirow[c]{5}{3cm}{\textit{Panel B: Census Tract Characteristics}} & Bachelor's degree, 2010 & 0.07 & 0.63 \\
 & Job density, 2013 & 0.00 & 0.70 \\
 & Median household income, 2016 & 0.49 & 0.03 \\
 & Poverty rate, 2010 & 0.00 & 0.00 \\
 & Population density, 2010 & 0.00 & 0.86 \\
\cline{1-4}
\multirow[c]{3}{3cm}{\textit{Panel C: Case Initiation}} & Filing for cause & 0.21 & 0.00 \\
 & Filing without cause & 0.01 & 0.03 \\
 & Filing for nonpayment & 0.00 & 0.00 \\
\cline{1-4}
\multirow[c]{4}{3cm}{\textit{Panel D: Defendant and Plaintiff Characteristics}} & Defendant has attorney & 0.02 & 0.00 \\
 & Plaintiff has attorney & 0.25 & 0.00 \\
 & Defendant is entity & 0.00 & 0.15 \\
 & Plaintiff is entity & 0.00 & 0.00 \\
\cline{1-4}
\bottomrule
\end{tabular}