### 2023-04-12 output/group_1_crimes_500m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\multirow[c]{2}{.75cm}{\textit{Panel A}} & Total Group 1 Incidents, 2017 & 57.77 & -0.39 & 0.80 & -0.94 & 0.55 \\
 & $\Delta$ Group 1 Incidents, 2017-2019 & -13.98 & 0.17 & 0.83 & 0.24 & 0.76 \\
\cline{1-7}
\multirow[c]{5}{.75cm}{\textit{Panel B}} & Bachelor's degree, 2010 & 0.32 & 0.01 & 0.63 & -0.00 & 0.72 \\
 & Job density, 2013 & 16,161.03 & -845.94 & 0.70 & -398.30 & 0.85 \\
 & Median household income, 2016 & 47,553.06 & 2,781.21 & 0.03 & -208.23 & 0.78 \\
 & Poverty rate, 2010 & 0.28 & -0.02 & 0.00 & -0.01 & 0.16 \\
 & Population density, 2010 & 23,320.19 & -128.01 & 0.86 & -374.93 & 0.60 \\
\cline{1-7}
\bottomrule
\end{tabular}