### 2023-04-13 output/group_1_crimes_500m/tables/balance_table.tex

\begin{tabular}{llccccc}
\toprule
 &  & \textit{} & \multicolumn{4}{c}{\textit{Difference in Cases Won by Defendant}} \\
\cline{4-7}
\\
 &  & Cases Won by Plaintiff & Unweighted & \emph{p} & Weighted & \emph{p} \\
\midrule
\textit{Panel A} & Total Group 1 Incidents, 2017 & 81.75 & -2.95 & 0.50 & -2.42 & 0.58 \\
\cline{1-7}
\multirow[c]{5}{.75cm}{\textit{Panel B}} & Bachelor's degree, 2010 & 0.32 & 0.00 & 0.81 & -0.01 & 0.44 \\
 & Job density, 2013 & 16,161.03 & -1,272.66 & 0.55 & -535.86 & 0.80 \\
 & Median household income, 2016 & 47,553.06 & 2,997.87 & 0.01 & -1,006.45 & 0.32 \\
 & Poverty rate, 2010 & 0.28 & -0.02 & 0.00 & -0.01 & 0.21 \\
 & Population density, 2010 & 23,320.19 & -297.77 & 0.67 & -643.10 & 0.37 \\
\cline{1-7}
\textit{Panel C} & Filing for nonpayment & 0.72 & -0.11 & 0.00 & -0.03 & 0.08 \\
\cline{1-7}
\multirow[c]{2}{.75cm}{\textit{Panel D}} & Plaintiff has attorney & 0.88 & -0.05 & 0.00 & -0.03 & 0.05 \\
 & Plaintiff is entity & 0.81 & -0.08 & 0.00 & -0.03 & 0.08 \\
\cline{1-7}
\bottomrule
\end{tabular}