### 2023-04-14 output/final_paper/tables/summary_statistics.tex

\begin{tabular}{llcccc}
\toprule
 &  & Mean & S.D. & N \\
Panel & Variable &  &  &  \\
\midrule
\multirow[c]{2}{4cm}{\textit{Panel A: Pre-treatment Outcomes}} & Change in Crime Incidents, 2017-2019 & -152.24 & 217.71 & 1,689 \\
 & Total Crime Incidents, 2017 & 1,164.18 & 714.31 & 1,689 \\
\cline{1-5}
\multirow[c]{6}{4cm}{\textit{Panel B: Census Tract Characteristics}} & Bachelor's degree, 2010 & 0.32 & 0.22 & 1,689 \\
 & Job density, 2013 & 16,714.10 & 43,368.36 & 1,689 \\
 & Median household income, 2016 & 46,250.26 & 24,935.09 & 1,689 \\
 & Poverty rate, 2010 & 0.29 & 0.15 & 1,689 \\
 & Population density, 2010 & 23,449.59 & 14,401.77 & 1,689 \\
 & share_white2010 & 0.31 & 0.27 & 1,689 \\
\cline{1-5}
\multirow[c]{3}{4cm}{\textit{Panel C: Case Initiation}} & Filing for cause & 0.11 & 0.32 & 1,689 \\
 & Filing without cause & 0.04 & 0.20 & 1,689 \\
 & Filing for nonpayment & 0.76 & 0.43 & 1,689 \\
\cline{1-5}
\multirow[c]{4}{4cm}{\textit{Panel D: Defendant and Plaintiff Characteristics}} & Defendant has attorney & 0.03 & 0.18 & 1,689 \\
 & Plaintiff has attorney & 0.91 & 0.29 & 1,689 \\
 & Defendant is entity & 0.02 & 0.15 & 1,689 \\
 & Plaintiff is entity & 0.84 & 0.36 & 1,689 \\
\cline{1-5}
\multirow[c]{9}{4cm}{\textit{Panel E: Case Resolution}} & Case duration & 20.10 & 25.06 & 1,595 \\
 & Judgment by default & 0.46 & 0.50 & 1,689 \\
 & Case dismissed & 0.43 & 0.50 & 1,689 \\
 & Case heard & 0.05 & 0.21 & 1,689 \\
 & Execution issued & 0.30 & 0.46 & 1,689 \\
 & Money judgment & 1,488.32 & 3,429.81 & 1,689 \\
 & Execution levied & 0.06 & 0.24 & 1,689 \\
 & Execution requested & 0.03 & 0.16 & 1,689 \\
 & Execution status unknown & 0.00 & 0.00 & 1,689 \\
\cline{1-5}
\bottomrule
\end{tabular}