### 2023-08-02 output/final_paper/tables/summary_statistics.tex

\begin{tabular}{llcccc}
\toprule
 &  & Mean & S.D. & N \\
Panel & Variable &  &  &  \\
\midrule
\multirow[c]{2}{4cm}{\textit{Panel A: Pre-treatment Outcomes}} & Change in Crime Incidents, 2017-2019 & -150.97 & 216.45 & 1,909 \\
 & Total Crime Incidents, 2017 & 1,164.90 & 717.22 & 1,909 \\
\cline{1-5}
\multirow[c]{6}{4cm}{\textit{Panel B: Census Tract Characteristics}} & Bachelor's degree, 2010 & 0.32 & 0.22 & 1,909 \\
 & Job density, 2013 & 16,845.17 & 44,087.24 & 1,909 \\
 & Median household income, 2016 & 46,653.89 & 24,977.84 & 1,909 \\
 & Poverty rate, 2010 & 0.29 & 0.15 & 1,909 \\
 & Population density, 2010 & 23,451.78 & 14,298.74 & 1,909 \\
 & Share white, 2010 & 0.31 & 0.27 & 1,909 \\
\cline{1-5}
\multirow[c]{3}{4cm}{\textit{Panel C: Case Initiation}} & Filing for cause & 0.13 & 0.34 & 1,909 \\
 & Filing without cause & 0.05 & 0.22 & 1,909 \\
 & Filing for nonpayment & 0.74 & 0.44 & 1,909 \\
\cline{1-5}
\multirow[c]{4}{4cm}{\textit{Panel D: Defendant and Plaintiff Characteristics}} & Defendant has attorney & 0.05 & 0.22 & 1,909 \\
 & Plaintiff has attorney & 0.91 & 0.29 & 1,909 \\
 & Defendant is entity & 0.02 & 0.15 & 1,909 \\
 & Plaintiff is entity & 0.83 & 0.38 & 1,909 \\
\cline{1-5}
\multirow[c]{5}{4cm}{\textit{Panel E: Case Resolution}} & Case duration & 41.75 & 81.98 & 1,814 \\
 & Judgment by default & 0.46 & 0.50 & 1,909 \\
 & Case dismissed & 0.44 & 0.50 & 1,909 \\
 & Case heard & 0.05 & 0.22 & 1,909 \\
 & Money judgment & 1,748.60 & 4,708.19 & 1,909 \\
\cline{1-5}
\bottomrule
\end{tabular}