### 2023-09-14 output/summary_statistics.tex

\begin{tabular}{llcccc}
\toprule
 &  & Mean & Median & S.D. & N \\
Panel & Variable &  &  &  &  \\
\midrule
\multirow[c]{6}{4cm}{\textit{Panel A: Pre-treatment Outcomes}} & Change in all crimes, 2017-2019 & -0.54 & 0.00 & 4.26 & 6,060 \\
 & Change in assaults, 2017-2019 & 0.01 & 0.00 & 1.00 & 6,060 \\
 & Change in investigations, 2017-2019 & 0.05 & 0.00 & 1.04 & 6,060 \\
 & All crimes, 2019-01 & 5.16 & 4.00 & 5.77 & 6,060 \\
 & Assault crimes, 2019-01 & 0.41 & 0.00 & 0.80 & 6,060 \\
 & Investigations, 2019-01 & 0.45 & 0.00 & 0.82 & 6,060 \\
\cline{1-6}
\multirow[c]{5}{4cm}{\textit{Panel B: Census Tract Characteristics}} & frac\_coll\_plus2010 & 0.31 & 0.24 & 0.22 & 6,060 \\
 & job\_density\_2013 & 14,806.08 & 2,425.15 & 40,595.37 & 6,060 \\
 & med\_hhinc2016 & 47,188.87 & 41,146.00 & 24,703.65 & 6,060 \\
 & poor\_share2010 & 0.28 & 0.25 & 0.14 & 6,060 \\
 & popdensity2010 & 22,704.52 & 20,644.44 & 13,091.45 & 6,060 \\
\cline{1-6}
\multirow[c]{3}{4cm}{\textit{Panel C: Case Initiation}} & for\_cause & 0.17 & 0.00 & 0.37 & 6,060 \\
 & no\_cause & 0.05 & 0.00 & 0.22 & 6,060 \\
 & non\_payment & 0.78 & 1.00 & 0.42 & 6,060 \\
\cline{1-6}
\multirow[c]{4}{4cm}{\textit{Panel D: Defendant and Plaintiff Characteristics}} & hasAttyD & 0.08 & 0.00 & 0.27 & 6,060 \\
 & hasAttyP & 0.88 & 1.00 & 0.32 & 6,060 \\
 & isEntityD & 0.02 & 0.00 & 0.14 & 6,060 \\
 & isEntityP & 0.81 & 1.00 & 0.39 & 6,060 \\
\cline{1-6}
\multirow[c]{6}{4cm}{\textit{Panel E: Case Resolution}} & case\_duration & 58.17 & 19.00 & 88.96 & 5,500 \\
 & defaulted & 0.19 & 0.00 & 0.39 & 6,060 \\
 & dismissed & 0.27 & 0.00 & 0.44 & 6,060 \\
 & heard & 0.03 & 0.00 & 0.17 & 6,060 \\
 & judgment & 1,938.57 & 0.00 & 5,337.87 & 6,060 \\
 & mediated & 0.40 & 0.00 & 0.49 & 6,060 \\
\cline{1-6}
\bottomrule
\end{tabular}